---
title: "CMC analysis"
author: "Christer Sylvén"
date: "`r format(Sys.time(), '%y-%m-%d')`"
output: 
  html_document:
    self_contained: true
    highlight: tango
    df_print: paged
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true 
---

```{r setup, include=FALSE}
knitr::opts_knit$set(progress=TRUE,verbose=TRUE)
knitr::opts_chunk$set(message=FALSE, warning=FALSE, #result='hold',fig.width=unit(18,"cm"), fig.height =unit(11,"cm"))
result='hold',fig.width=12, fig.height =6)

```

Filtered count matrices for sqRNA and ST data from [Asp et al 2019](https://www.spatialresearch.org). Further Asp et al 2019 preprocessed matrices obtained through this workflow. Saved as: data2/all_cells_count_matrix_filtered.tsv (scRNA) and data2/filtered_matrix.tsv (STdata). For ST maps cardiac background data file grobs.list is needed. Please contact authors.

# UMAP of Asp et al 2019 data

```{r}
#if (!requireNamespace("BiocManager", quietly = TRUE))
#  install.packages("BiocManager")

#BiocManager::install("GSEABase")
#BiocManager::install("AUCell")
#BiocManager::install("doRNG")
#BiocManager::install("org.Hs.eg.db")

suppressPackageStartupMessages({
  library(data.table)
  library(Seurat)
  library(magrittr)
  library(dplyr)
  # library(biomaRt)
  library(ggplot2)
  library(cowplot)
  library(plotly)
  #library(DT)
  # library(IRdisplay)
  library(grid)
  library(org.Hs.eg.db)
  library(clusterProfiler)
  library(GSEABase)
  library(AUCell)
  library(doRNG)
  # library(repr)
  # library(ggrepel)
  # library(sctransform)
  # library(gridExtra)
  library(openxlsx)
  library(reshape2)
  library(readxl)
  library(forcats)
  library(patchwork)
  library(corrplot)
# library(biomaRt)
})

path1 <-'/Users/ChristerSylven/Human-Prenatal-Cardiomyocytes/'
extractTopN <- function(DE_genes, N = 5) {
  
  subset(DE_genes, avg_log2FC > 0) %>% # Subset the data.frame to include only positive avg_logFC values
    arrange(-avg_log2FC) %>% # Arrange the data.frame by decresing avg_logFC ("-" = decreasing)
    group_by(cluster) %>% # Group the data.frame by cluster (this will produce a "grouped" data.frame, i.e. a subset for each cluster)
    top_n(N, avg_log2FC) %>% # Extract top N rows for each cluster based on avg_logFC
    arrange(cluster) 
}

# original filtered count matrices from Asp et al 2019: Count matrices are available at https://www.spatialresearch.org. Reporting 3777 cells

seu_0 <- read.table(file = paste0(path1,'data2/all_cells_count_matrix_filtered.tsv'),  sep = '\t', header = TRUE, stringsAsFactors=FALSE)
rownames(seu_0) <- seu_0[,1]
seu_0 <- seu_0[,-1]

# 3717 cells used in Asp et al using Seurat 2.3
Article_rownames <- readRDS(paste0(path1, 'data/Article_rownames_SC.R'))
seu<- seu_0[,Article_rownames]
dim(seu)
seu <- CreateSeuratObject(counts = seu, project = "seu_SC", min.cells = 3, min.features = 200)
#seu
```

### S_Figure 1

```{r}
# store mitochondrial percentage in object meta data
seu[["percent.mt"]] <- PercentageFeatureSet(seu, pattern = "^MT-")
seu <- subset(seu, subset = nFeature_RNA > 200 & nFeature_RNA < 6000 & nCount_RNA < 50000 & percent.mt < 20)
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
seu <- CellCycleScoring(seu, s.features = s.genes, g2m.features = g2m.genes,  set.ident = TRUE )
seu <- SCTransform(seu,   assay = 'RNA',   new.assay.name = 'SCT',   vars.to.regress = c('percent.mt', 'S.Score', 'G2M.Score'), verbose = FALSE )
seu <- RunPCA(seu, pcs = 30, verbose = FALSE)
seu <- FindNeighbors(seu, dims = 1:30, verbose = FALSE)
seu <- FindClusters(seu, resolution = 0.38, verbose = FALSE)
#seu <- RunTSNE(seu, dims = 1:30,  tsne.method = "Rtsne")
seu <- RunUMAP(seu, dims = 1:30, verbose = FALSE)
#p1 <- DimPlot(seu, reduction = "tsne", pt.size = 0.25, cols = fill2, label='TRUE', repel='TRUE') + ggtitle(label = "t-SNE")
fill2 = c('#1F77B4', '#AEC7E8', '#FF7F0E', '#FFBB78', '#2CA02C',  'red', '#F7B6D2', '#9467BD', '#D62728',  '#C5B0D5', '#8C564B', '#C49C94', 'green', '#E377C2',  '#BCBD22',  '#17BECF','#DBDB8D')

# To get text annotation to add to UMAP figure
#Asp et al 2019, modified by Seurat 3.0
new.cluster.ids <- c(" 0  Endocardium/endothelium", #0
                     " 1  Ventricular cardiomyocytes", #1 
                     " 2  Epicardium-derived cells",
                     " 3  Fibroblast-like (cardiac skeleton)",
                     " 4  Fibroblast-like (smaller vascular development)",
                     " 5  Erythrocytes", 
                     " 6  Fibroblast-like (larger vascular development)",
                     " 7  Fibroblast-like smooth muscle cells",
                     " 8  Atrial cardiomyocytes",
                     " 9  Endothel/pericytes/adventitia",
                     "10  Smooth muscle cells/fibroblast-like",
                     "11  Erythrocytes",
                     "12  Epicardium",
                     "13  MYOZ2-enriched cardiomyocytes",
                     "14  Immune cells",
                     "15  Schwann progenitor cells",
                     "16  Cardiac neural crest cells")
```

```{r}
p2 <- DimPlot(seu, reduction = "umap",  cols = alpha(fill2, .60), label.size = 6, pt.size = 1.5,  label='TRUE') + scale_color_manual(labels = new.cluster.ids, values = fill2) + labs(color = "")
p2

```

# Umap of cardiomyocyte subclusters

after subsetting the 3 original cardiomyocyte clusters analysed as new Seurat object with identical pipe.

```{r}
# analyse cardiomyocyte clusters with same setup. 
Idents(seu) <- 'SCT_snn_res.0.38'
seu_Muscle <- subset(seu, idents=c(1,8,13))
#seu_Muscle
#names(seu_Muscle[[]])
# eliminate base cluster names w zero cells
base_clusters  <- factor(seu_Muscle@meta.data[,11])
#length(base_clusters)
seu_Muscle <- seu_0[,rownames(seu_Muscle@meta.data)] # raw data from original counts file

dim(seu_Muscle)

```

### Figure 1A, pcs = 30

```{r}
seu_Muscle <- CreateSeuratObject(counts = seu_Muscle, project = "seu_Muscle_SC", min.cells = 3, min.features = 200)
#seu_Muscle
# store mitochondrial percentage in object meta data
seu_Muscle <- PercentageFeatureSet(seu_Muscle, pattern = "^MT-", col.name = "percent.mt")
seu_Muscle <- subset(seu_Muscle, subset = nFeature_RNA > 200 & nFeature_RNA < 6000 & nCount_RNA < 50000 & percent.mt < 20)
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
seu_Muscle <- CellCycleScoring(seu_Muscle, s.features = s.genes, g2m.features = g2m.genes,  set.ident = TRUE )
seu_Muscle <- SCTransform(seu_Muscle,   assay = 'RNA',   new.assay.name = 'SCT',   vars.to.regress = c('percent.mt', 'S.Score', 'G2M.Score'), verbose = FALSE )
seu_Muscle <- RunPCA(seu_Muscle, pcs = 39, verbose = FALSE)
seu_Muscle <- FindNeighbors(seu_Muscle, dims = 1:30, verbose = FALSE)
seu_Muscle <- FindClusters(seu_Muscle, resolution = 0.7, verbose = FALSE)
#seu_Muscle <- RunTSNE(seu_Muscle, dims = 1:30,  tsne.method = "Rtsne")
seu_Muscle <- RunUMAP(seu_Muscle, dims = 1:30, verbose = FALSE)
#DimPlot(seu_Muscle, reduction = "umap", pt.size = 0.25,  label.size =6, cols = fill2, label='TRUE') + ggtitle(label = "")

seu_Muscle@meta.data$ClusterNames <- seu_Muscle@meta.data[,11]
seu_Muscle@meta.data[,13] <- as.character(seu_Muscle@meta.data[,13])
seu_Muscle@meta.data[,13][which(seu_Muscle@meta.data[,13]=='0')] <- 'Trabecular'
seu_Muscle@meta.data[,13][which(seu_Muscle@meta.data[,13]=='1')] <- 'Compact'
seu_Muscle@meta.data[,13][which(seu_Muscle@meta.data[,13]=='2')] <- 'Purkinje-related'
seu_Muscle@meta.data[,13][which(seu_Muscle@meta.data[,13]=='3')] <- 'Atrial trabecular'
seu_Muscle@meta.data[,13][which(seu_Muscle@meta.data[,13]=='4')] <- 'High G2M'
seu_Muscle@meta.data[,13][which(seu_Muscle@meta.data[,13]=='5')] <- 'Atrial conduit'
seu_Muscle@meta.data[,13][which(seu_Muscle@meta.data[,13]=='6')] <- 'Exosome-related'
seu_Muscle@meta.data[,13][which(seu_Muscle@meta.data[,13]=='7')] <- 'Outflow tract'
#seu_Muscle@meta.data[,13][which(seu_Muscle@meta.data[,13]=='8')] <- 'SAN'
seu_Muscle@meta.data[,13] <- as.factor(seu_Muscle@meta.data[,13])

seu_Muscle$base_clusters <- base_clusters


fill4=c('darkmagenta', 'royalblue1', '#FF7F0E', 'maroon', 'blue','lightgreen', 'green',  'magenta', 'red','yellow',  'grey')

# in order largest to smallest cluster
seu_Muscle@meta.data$ClusterNames <- factor(seu_Muscle@meta.data$ClusterNames, levels = c('Trabecular','Compact',  'Purkinje-related','Atrial trabecular', 'High G2M',  'Atrial conduit','Exosome-related', 'Outflow tract'))

Idents(seu_Muscle) <- 'ClusterNames'
DimPlot(seu_Muscle, reduction = "umap", cols = alpha(fill4, 0.60), pt.size = 2.0, label='F')

D1 <- DimPlot(seu_Muscle, reduction = "umap", cols = alpha(fill4, 0.60), pt.size = 2.0, label='F')  + ggtitle('pcs = 30') &  NoAxes()  &  NoLegend()

```

```{r}
seu_Muscle <- RunPCA(seu_Muscle, pcs = 25, verbose = FALSE)
seu_Muscle <- FindNeighbors(seu_Muscle, dims = 1:30, verbose = FALSE)
seu_Muscle <- FindClusters(seu_Muscle, resolution = 0.75, verbose = FALSE)
seu_Muscle <- RunUMAP(seu_Muscle, dims = 1:30, verbose = FALSE)
D2 <- DimPlot(seu_Muscle, reduction = "umap", cols = alpha(fill4, 0.60), pt.size = 2.0, label='F')  + ggtitle('pcs = 25') &  NoAxes()  &  NoLegend()
```


### Figure S 1B_pcs_25_30_35

```{r}
seu_Muscle <- RunPCA(seu_Muscle, pcs = 35, verbose = FALSE)
seu_Muscle <- FindNeighbors(seu_Muscle, dims = 1:30, verbose = FALSE)
seu_Muscle <- FindClusters(seu_Muscle, resolution = 0.75, verbose = FALSE)
#seu_Muscle <- RunTSNE(seu_Muscle, dims = 1:30,  tsne.method = "Rtsne")
seu_Muscle <- RunUMAP(seu_Muscle, dims = 1:30, verbose = FALSE)

Idents(seu_Muscle) <- 'ClusterNames'
D3 <- DimPlot(seu_Muscle, reduction = "umap", cols = alpha(fill4, 0.60), pt.size = 2.0, label='F')  + ggtitle('pcs = 35') & NoAxes()
D2 + D1 + D3


```

### Figure 1B

```{r fig1B, fig.height = 9, fig.width = 18, fig.align = "center" }

# calculate number of RP genes in the different clusters
Idents(seu_Muscle) <- 'ClusterNames'

seu_Muscle.markers <- FindAllMarkers(seu_Muscle, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)


sign <- seu_Muscle.markers[which(seu_Muscle.markers$p_val_adj<0.05),]
table(sign$cluster)
sign$ribo <- substr(sign$gene,1,2) # two first letters of gene

sign$ribo[sign$ribo !='RP'] <- '0'
table(sign$cluster, sign$ribo)

sign6 <- sign[which(sign$cluster =='Exosome-related'),]
sign6 <- sign6[which(sign6$ribo !='RP'),] # 38 RPL (60S ribosomal subunit)18 RPS (40S ribosomal subunit)
#sign6 # genes that are not rpl or rps



muscle <- data.frame(group = c('Trabecular', 'Compact', 'Purkinje-related', 'High G2M', 'Exosome-related', 'Outflow tract'),
                   Percent = c(32.3, 24.0, 17.5, 12.6, 9.7, 3.9), #
                   Genes = c(3.2, 2.8, 2.6,  2.4, 1.5, 2.0),
                   #Genes_sd = c(0.7, 0.6, 0.6,  0.9, 0.6, 0.2),
                   Genes_sem = c(0.05, 0.05, 0.06,  0.10, 0.08, 0.25), #SEM
                   Transcripts = c(14.7, 11.9, 9.2,  9.6, 6.1, 7.9),
                  # Transcripts_sd = c(5.7, 3.6, 2.9,  4.7, 3.3, 6.6),
                   Transcripts_sem = c(0.4, 0.3, 0.3,  0.5, 0.4, 1.4),  #SEM
                   G2M_phase = c(6.8, 9.2, 4.9,  66.2, 5.3, 30.4),
                   S_phase = c(15.3, 16.3, 9.7,  5.4, 12.3, 21.7),
                   Percent_mito = c(8.8, 9.2, 8.5,  5.8, 1.3, 1.5),
                   #Percent_mito_sd = c(3.5, 3.0, 3.7,  5.1, 3.2, 2.4),
                   Percent_mito_sem = c(0.3, 0.3, 0.4,  0.6, 0.4, 0.5), #SEM
                   MYH7 = c(78, 75, 53,  38, 14, 7),
                   #MYH7_sd = c(32, 25, 27,  31, 12, 14),
                   MYH7_sem = c(2.3, 2.1, 2.7,  3.7, 1.6, 2.9),#SEM
                   GJA1_Cx43 = c(1.5, 1.6, 2.2,  1.2, 0.3, 0.1),
                   #GJA1_Cx43_sd = c(1.9, 1.8, 2.3,  1.9, 0.7, 0.3),
                   GJA1_Cx43_sem = c(0.1, 0.2, 0.2,  0.2, 0.09, 0.06), #SEM
                   GJC1_Cx45 = c(0.5, 0.5, 0.8,  0.3, 0.1, 0.2),
                   #GJC1_Cx45_sd = c(0.7, 0.7, 1.0,  0.5, 0.4, 0.7),
                   GJC1_Cx45_sem = c(0.05, 0.06, 0.1,  0.06, 0.05, 0.15), #SEM
                   L_R_interactions  = c(1309, 1126, 1364,  704, 198, 182),
                   RP = c(0, 9.16, 0, 1.77, 68.29, 20.93)
                              )


muscle$group2 <- factor(muscle$group, levels = c( 'Trabecular', 'Compact', 'Purkinje-related', 'High G2M', 'Exosome-related', 'Outflow tract'))

fill_ventricle2 = c('darkmagenta','royalblue1',  '#FF7F0E', 'blue', 'green', 'magenta')

percent <- ggplot(muscle, aes(group2, Percent, fill = group2)) + geom_col(position = "dodge") +
  xlab('') +
  scale_y_continuous(position = "left", limits = c(0, 40), expand = c(0, 0)) +
  scale_fill_manual(values = fill_ventricle2) +
  ylab("Cardiomyocyte clusters (%)") +
  scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "") +
 # theme(legend.position="none", axis.text = element_text(size = 16),  axis.text.x = element_blank()) +
  theme_bw() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none", text = element_text(size = 20), axis.text = element_text(size = 20),  axis.text.x = element_blank(), axis.line.x = element_line(color = "white"), axis.line.y = element_line(color = "black"))

genes <- ggplot(muscle, aes(group2, Genes, fill = group2)) + geom_col(position = "dodge") +
  geom_errorbar(aes(ymin=Genes-Genes_sem,ymax=Genes+Genes_sem),width=.2,position=position_dodge(.9)) +
  xlab('') +
  scale_y_continuous(position = "left", limits = c(0, 4), expand = c(0, 0)) +
  scale_fill_manual(values = fill_ventricle2) +
  labs(title = expression(underline("****"))) + 
  ylab("Genes") +
  scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "") +
  theme_bw() +
  theme(plot.title= element_text(hjust=0.5, vjust = -3, size=rel(1.5)), panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position="none", text = element_text(size = 20), axis.text = element_text(size = 20), axis.line.x = element_line(color = "white"), axis.line.y = element_line(color = "black")) 


transcripts <- ggplot(muscle, aes(group2, Transcripts, fill = group2)) + geom_col(position = "dodge") +
  geom_errorbar(aes(ymin=Transcripts-Transcripts_sem,ymax=Transcripts+Transcripts_sem),width=.2,position=position_dodge(.9)) +
  xlab('') +
  scale_y_continuous(position = "left", limits = c(0, 80), expand = c(0, 0)) +
  scale_fill_manual(values = fill_ventricle2) +
  labs(title = expression(underline("****"))) + 
  scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "") +
  theme(plot.title= element_text(hjust=0.5, vjust = -3, size=rel(1.5)), panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position="none", text = element_text(size = 20), axis.text = element_text(size = 20), axis.line.x = element_line(color = "white"), axis.line.y = element_line(color = "black")) 

g2mphase <- ggplot(muscle, aes(group2, G2M_phase, fill = group2)) + geom_col(position = "dodge") +
  xlab('') +
  scale_y_continuous(position = "left", limits = c(0, 80), expand = c(0, 0)) +
  scale_fill_manual(values = fill_ventricle2) +
  ylab("G2M-phase (%)") +
  scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "") +
  labs(title = expression(underline("****"))) + 
  theme_bw() +
  theme(plot.title= element_text(hjust=0.5, vjust = -3, size=rel(1.5)), panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position="none", text = element_text(size = 20), axis.text = element_text(size = 20), axis.line.x = element_line(color = "white"), axis.line.y = element_line(color = "black")) 


sphase <- ggplot(muscle, aes(group2, S_phase, fill = group2)) + geom_col(position = "dodge") +
  xlab('') +
  scale_y_continuous(position = "left", limits = c(0, 25), expand = c(0, 0)) +
  scale_fill_manual(values = fill_ventricle2) +
  ylab("S-phase (%)") +
  scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "") +
  labs(title = expression(underline("****"))) + 
  theme_bw() +
  theme(plot.title= element_text(hjust=0.5, vjust = -3, size=rel(1.5)), panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position="none", text = element_text(size = 20), axis.text = element_text(size = 20), axis.line.x = element_line(color = "white"), axis.line.y = element_line(color = "black")) 

percentmito <- ggplot(muscle, aes(group2, Percent_mito, fill = group2)) + geom_col(position = "dodge") +
  geom_errorbar(aes(ymin=Percent_mito-Percent_mito_sem,ymax=Percent_mito+Percent_mito_sem),width=.2,position=position_dodge(.9)) +
  xlab('') +
  scale_y_continuous(position = "left", limits = c(0, 10), expand = c(0, 0)) +
  scale_fill_manual(values = fill_ventricle2) +
  ylab("Mito (%)") +
  scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "") +
  labs(title = expression(underline("****"))) + 
  theme_bw() +
  theme(plot.title= element_text(hjust=0.5, vjust = -3, size=rel(1.5)), panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position="none", text = element_text(size = 20), axis.text = element_text(size = 20), axis.line.x = element_line(color = "white"), axis.line.y = element_line(color = "black")) 


myh7 <- ggplot(muscle, aes(group2, MYH7, fill = group2)) + geom_col(position = "dodge") +
  geom_errorbar(aes(ymin=MYH7-MYH7_sem,ymax=MYH7+MYH7_sem),width=.2,position=position_dodge(.9)) +
  xlab('') +
  scale_y_continuous(position = "left", limits = c(0, 100), expand = c(0, 0)) +
  scale_fill_manual(values = fill_ventricle2) +
  ylab("MYH7 (counts)") +
  scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "") +
  labs(title = expression(underline("****"))) + 
  theme_bw() +
  theme(plot.title= element_text(hjust=0.5, vjust = -3, size=rel(1.5)), panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position="none", text = element_text(size = 20), axis.text = element_text(size = 20), axis.line.x = element_line(color = "white"), axis.line.y = element_line(color = "black")) 

gja1 <- ggplot(muscle, aes(group2, GJA1_Cx43, fill = group2)) + geom_col(position = "dodge") +
  geom_errorbar(aes(ymin=GJA1_Cx43-GJA1_Cx43_sem,ymax=GJA1_Cx43+GJA1_Cx43_sem),width=.2,position=position_dodge(.9)) +
  xlab('') +
  scale_y_continuous(position = "left", limits = c(0, 3), expand = c(0, 0)) +
  scale_fill_manual(values = fill_ventricle2) +
  ylab("GJA1-Cx43 (counts)") +
  scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "") +
  labs(title = expression(underline("****"))) + 
  theme_bw() +
  theme(plot.title= element_text(hjust=0.5, vjust = -3, size=rel(1.5)), panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position="none", text = element_text(size = 20), axis.text = element_text(size = 20), axis.line.x = element_line(color = "white"), axis.line.y = element_line(color = "black"))

gja1_legend <- ggplot(muscle, aes(group2, GJA1_Cx43, fill = group2)) + geom_col(position = "dodge") +
  geom_errorbar(aes(ymin=GJA1_Cx43-GJA1_Cx43_sem,ymax=GJA1_Cx43+GJA1_Cx43_sem),width=.2,position=position_dodge(.9)) +
  xlab('') +
  scale_y_continuous(position = "left") +
  scale_fill_manual(values = fill_ventricle2) +
  labs(title="****") + 
  theme( plot.title = element_text(hjust=0.5, size=rel(1.5)), text = element_text(size = 20), axis.text = element_text(size = 20), legend.text = element_text(size =20),legend.title=element_blank())

#if(!require(devtools)) install.packages("devtools")
#devtools::install_github("kassambara/ggpubr")
library(ggpubr)

# Extract the legend. Returns a gtable
leg <- get_legend(gja1_legend)
legend <-  as_ggplot(leg)


gjc1 <- ggplot(muscle, aes(group2,  GJC1_Cx45, fill = group2)) + geom_col(position = "dodge") +
  geom_errorbar(aes(ymin= GJC1_Cx45- GJC1_Cx45_sem,ymax= GJC1_Cx45+ GJC1_Cx45_sem),width=.2,position=position_dodge(.9)) +
  xlab('') +
  scale_y_continuous(position = "left", limits = c(0, 1), expand = c(0, 0)) +
  scale_fill_manual(values = fill_ventricle2) +
  ylab("GJC1-Cx45 (counts)") +
  scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "") +
  labs(title = expression(underline("****"))) + 
  theme_bw() +
  theme(plot.title= element_text(hjust=0.5, vjust = -3, size=rel(1.5)), panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position="none",text = element_text(size = 20),  axis.text = element_text(size = 20), axis.line.x = element_line(color = "white"), axis.line.y = element_line(color = "black"))

l_r<- ggplot(muscle, aes(group2, L_R_interactions, fill = group2)) + geom_col(position = "dodge") +
  xlab('') +
  scale_y_continuous(position = "left", limits = c(0, 1500), expand = c(0, 0)) +
  scale_fill_manual(values = fill_ventricle2) +
  ylab("Ligand-receptor interactions") +
  scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "") +
  labs(title = expression(underline("****"))) + 
  theme_bw() +
  theme(plot.title= element_text(hjust=0.5, vjust = -3, size=rel(1.5)), panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position="none", text = element_text(size = 20), axis.text = element_text(size = 20), axis.line.x = element_line(color = "white"), axis.line.y = element_line(color = "black"))


rp<- ggplot(muscle, aes(group2, RP, fill = group2)) + geom_col(position = "dodge") +
  xlab('') +
  scale_y_continuous(position = "left", limits = c(0, 80), expand = c(0, 0)) +
  scale_fill_manual(values = fill_ventricle2) +
  ylab("RP (%) of DE genes") +
  scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "") +
  labs(title = expression(underline("****"))) + 
  theme_bw() +
  theme(plot.title= element_text(hjust=0.5, vjust = -3, size=rel(1.5), face = 'bold'), panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  legend.position="none", text = element_text(size = 20), axis.text = element_text(size = 20), axis.line.x = element_line(color = "white"), axis.line.y = element_line(color = "black"))  
                                                                                                                                      


percent + genes  + myh7 + gja1 + gjc1 + percentmito + g2mphase + sphase + rp + l_r + legend + plot_layout(widths = c(1,1,1,1,1,1,1,1,1,1,1),  ncol = 6) #width 12, height 6


```

### Subsetting atrial clusters shown in UMAP and FeaturePlots

Identifying SAN cells as defined as coexpression of TBX18, SHOX2 and HCN4 according to literature.

```{r}
# identify SAN cluster as cells coexpressing SHOX2, TBX18 and HCN4
GOI.1 <- which(GetAssayData(object= seu_Muscle, slot = 'data')['SHOX2',] > 0 & GetAssayData(object=seu_Muscle, slot = 'data')['TBX18',] > 0 & GetAssayData(object=seu_Muscle, slot = 'data')['HCN4',] > 0)
length(GOI.1)

# SAN3 n=8
seu_Muscle@meta.data$sub_cluster_SAN <- seu_Muscle@meta.data[,11]
seu_Muscle@meta.data$sub_cluster_SAN <- as.character(seu_Muscle@meta.data$sub_cluster_SAN)
seu_Muscle@meta.data[c(names(GOI.1)), 16] <- '8'
seu_Muscle@meta.data$sub_cluster_SAN <- factor(seu_Muscle@meta.data$sub_cluster_SAN,  levels = c('0','1',  '2','3', '4',  '5','6', '7', '8'))

# create seu metadata file with muscle subclusters
#names(seu[[]])
#names(seu_Muscle[[]])
#seu
#seu_Muscle
#table(seu@meta.data[,12])
#table(seu_Muscle@meta.data[,15])
seu_Muscle@meta.data$sub_cluster_SAN2 <- seu_Muscle@meta.data[,16]
seu_Muscle@meta.data[,17] <- as.character(seu_Muscle@meta.data[,17])
seu_Muscle@meta.data[,17][which(seu_Muscle@meta.data[,17]=='0')] <- 'M0'
seu_Muscle@meta.data[,17][which(seu_Muscle@meta.data[,17]=='1')] <- 'M1'
seu_Muscle@meta.data[,17][which(seu_Muscle@meta.data[,17]=='2')] <- 'M2'
seu_Muscle@meta.data[,17][which(seu_Muscle@meta.data[,17]=='3')] <- 'M3'
seu_Muscle@meta.data[,17][which(seu_Muscle@meta.data[,17]=='4')] <- 'M4'
seu_Muscle@meta.data[,17][which(seu_Muscle@meta.data[,17]=='5')] <- 'M5'
seu_Muscle@meta.data[,17][which(seu_Muscle@meta.data[,17]=='6')] <- 'M6'
seu_Muscle@meta.data[,17][which(seu_Muscle@meta.data[,17]=='7')] <- 'M7'
seu_Muscle@meta.data[,17][which(seu_Muscle@meta.data[,17]=='8')] <- 'SAN'
seu_Muscle@meta.data[,17] <- as.factor(seu_Muscle@meta.data[,17])

seu_merge <- seu@meta.data[,11, drop = FALSE]
seu_Muscle_merge <- seu_Muscle@meta.data[,17, drop = FALSE]
names(seu_merge)[1] <- 'sub_cluster_SAN2'
seu_merge$row <- rownames(seu_merge)
seu_Muscle_merge$row <- rownames(seu_Muscle_merge)

t <- seu_merge %>%
  left_join(seu_Muscle_merge, by = "row") %>% 
  mutate(sub_cluster_SAN2 = coalesce(sub_cluster_SAN2.y, sub_cluster_SAN2.x))
t <- t[,c(2,4)]
t[,2] <- factor(t[,2])
#dim(t)
seu@meta.data$sub_cluster_SAN2 <- t[,2]
Idents(seu) <- 'sub_cluster_SAN2'
#DimPlot(seu, reduction = 'umap', label =T)

saveRDS(seu, file = paste0(path1, 'seu_2021_May.R'))
saveRDS(seu_Muscle, file = paste0(path1, 'seu_Muscle_2021_May.R'))

#seu <- readRDS(paste0(path1, 'seu_2021_May.R'))
Idents(seu) <- 'SCT_snn_res.0.38'
seu_Muscle <- readRDS(paste0(path1,'seu_Muscle_2021_May.R'))
#names(seu_Muscle[[]])
#Idents(seu_Muscle) <- "SCT_snn_res.0.7" # original in Seurat
#Idents(seu_Muscle) <- "sub_cluster_SAN" # with SAN added
#table(seu_Muscle@meta.data[,15])
```

### Figure 2A

```{r figs2A, fig.height = 3, fig.width = 8, fig.align = "center" }
#Atria
Idents(seu_Muscle) <- 'sub_cluster_SAN'
Atria <- subset(seu_Muscle, idents = c('3', '5', '8'))

Atria@meta.data$sub_cluster_SAN <- factor(Atria@meta.data$sub_cluster_SAN)
Atria@meta.data$sub_cluster_SAN2 <- as.character(Atria@meta.data$sub_cluster_SAN)
Atria@meta.data[,17][which(Atria@meta.data[,16]=='3')] <- 'Trabecular'
Atria@meta.data[,17][which(Atria@meta.data[,16]=='5')] <- 'Conduit'
Atria@meta.data[,17][which(Atria@meta.data[,16]=='8')] <- 'SAN'
saveRDS(Atria, file = paste0(path1, 'Atria_July21.R'))
#Atria <- readRDS(paste0(path1, 'Atria_July21.R'))
Idents(Atria) <- 'sub_cluster_SAN2'
p1 <- DimPlot(Atria, cols = alpha(c('blue','green', 'red'),0.60), pt.size = 2 ) + theme(legend.justification = "top") & NoAxes() 

p2 <- FeaturePlot(Atria,  features = c("TBX18", "SHOX2", "HCN4",  'PPP1R1B', 'ISL1', 'TBX3',  "NPPA", 'NKX2-5', "MYH6",  "GJA5"  ), cols = c("yellow", "red"),  order = TRUE) & NoLegend() & NoAxes() 

p21 <- FeaturePlot(Atria,  features = "TBX18", cols = c("yellow", "red"),  order = TRUE) +  theme_gray() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  & NoLegend() & NoAxes() 
p22 <- FeaturePlot(Atria,  features = "SHOX2", cols = c("yellow", "red"),  order = TRUE) +  theme_gray() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  & NoLegend() & NoAxes() 
p23 <- FeaturePlot(Atria,  features = "HCN4", cols = c("yellow", "red"),  order = TRUE)  +  theme_gray() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  & NoLegend() & NoAxes() 
p24 <- FeaturePlot(Atria,  features = "PPP1R1B", cols = c("yellow", "red"),  order = TRUE) +  theme_gray() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  & NoLegend() & NoAxes() 
p25 <- FeaturePlot(Atria,  features = "ISL1", cols = c("yellow", "red"),  order = TRUE) +  theme_gray() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  & NoLegend() & NoAxes() 
p26 <- FeaturePlot(Atria,  features = "TBX3", cols = c("yellow", "red"),  order = TRUE) +  theme_gray() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  & NoLegend() & NoAxes() 
p27 <- FeaturePlot(Atria,  features = "NPPA", cols = c("yellow", "red"),  order = TRUE) +  theme_gray() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  & NoLegend() & NoAxes() 
p28 <- FeaturePlot(Atria,  features = "NKX2-5", cols = c("yellow", "red"),  order = TRUE) +  theme_gray() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  & NoLegend() & NoAxes() 
p29 <- FeaturePlot(Atria,  features = "MYH6", cols = c("yellow", "red"),  order = TRUE) +  theme_gray() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  & NoLegend() & NoAxes() 
p210 <- FeaturePlot(Atria,  features = "GJA5", cols = c("yellow", "red"),  order = TRUE) +  theme_gray() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  & NoLegend() & NoAxes() 

p2 <- p21+p22+p23+p24+p25+p26+p27+p28+p29+p210 + plot_layout( ncol = 5)

p1 + p2  + plot_layout(widths = c(1,3))
```

## DotPlot for ventricular DE genes

### S_Figure 2A

```{r}
Idents(seu_Muscle) <- 'sub_cluster_SAN'
seu_Muscle.markers <- FindAllMarkers(seu_Muscle, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
#dim(seu_Muscle.markers)
#extract markers for dotplot & AUC
#DotPLot for Ventricles
markers_Muscle <- extractTopN(seu_Muscle.markers, N = 12) # for Ventricle n = 12, 


options(repr.plot.width = 10, repr.plot.height = 20)




markers_Muscle <- data.frame(markers_Muscle[,c(6:7, 5, 2)])
#markers_Muscle
markers_Muscle.to.plot <- unique(markers_Muscle$gene)

Ventricle <- c(0:2,4,6:7)
markers_Muscle_Ventricle <- filter(markers_Muscle, cluster %in%  Ventricle) # Atria clusters deleted
markers_Muscle_Ventricle.to.plot <- unique(markers_Muscle_Ventricle$gene)

seu_Ventricle<- subset(seu_Muscle, idents = c(0:2,4,6:7))

heat_cols <- rev(c("red", 'grey80', 'grey95')) # base: cols = c('blue', 'white') in DotPlot

d <- DotPlot(object = seu_Ventricle,  features = markers_Muscle_Ventricle.to.plot, dot.scale = 4, group.by ='sub_cluster_SAN') + 
  #scale_colour_gradientn(colors = heat_cols) +
  scale_y_discrete(name  = "", 
                   labels = c("Trabecular", "Compact", "Purkinje-related",
                              "High G2M",
                              "Exosome-related",
                              "Outflow tract")) +
  theme(legend.position= 'top', text = element_text(size = 12),axis.text.x = element_text(size = 8, angle = -90, hjust = 0,vjust = 1), axis.text.y = element_text(size = 10), plot.margin = unit(c(0.5,1,0,0), "cm")) +
  xlab('  ')

d +
  annotate ("rect", xmin = 0, xmax = 12.5, ymin = 0, ymax = 6.5, alpha = .1, fill = "blue") +
  annotate ("rect", xmin = 23.5, xmax = 35.5, ymin = 0, ymax = 6.5, alpha = .1, fill = "blue") +
  annotate ("rect", xmin = 47.5, xmax = 58.5, ymin = 0, ymax = 6.5, alpha = .1, fill = "blue") 


#ggsave("FiguresVentricle_dotplot_0602.pdf", d, height = 15, width = 40,  device = "pdf", units = "cm")
#dev.off()

```

### Heatmap

```{r fig heat_ventricle, fig.height = 7, fig.width = 7, fig.align = "center"}

Idents(seu_Ventricle) <- 'ClusterNames'

d <- DoHeatmap(object = seu_Ventricle,  features = markers_Muscle_Ventricle.to.plot, angle = 90, size = 2.5) + NoLegend() + 
    theme(text = element_text(size = 8))
#ggsave(d, filename = paste0(path1,"Figures/heatmap_ventricle.svg",  width = 12.5, height = 25,  device = "svg", units = "cm")
#dev.off()

d

```



### Figure S3A

### DE genes for atria with SHOX2, TBX18 and HCN4 removed

```{r}
library(Seurat)


 counts2<- GetAssayData(seu_Muscle, assay = "RNA")
  counts2 <- counts2[-(which(rownames(counts2) %in% c('SHOX2','TBX18', 'HCN4'))),]
  Atrium_minus <- subset(seu_Muscle, features = rownames(counts2))
  
  Idents(Atrium_minus) <- 'sub_cluster_SAN'
 Atrium_minus.markers <- FindAllMarkers(Atrium_minus, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

markers_Atrium_minus  <- extractTopN(Atrium_minus.markers, N = 18) # for Atria n = 18
markers_Atrium_minus$cluster <- as.character(markers_Atrium_minus$cluster) 

markers_Atrium_minus  <- data.frame(markers_Atrium_minus[,c(6:7, 5, 2)])

#Atria <- readRDS('Atria_July21.R')
Atria2 <- c(3,5,8)

Atrium_minus2 <- subset(Atrium_minus, idents = Atria2)

 #Atrium_minus2.markers <- FindAllMarkers(Atrium_minus2, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
 #Atrium_minus2.markers  <- extractTopN(Atrium_minus2.markers, N = 18) # for Atria n = 18
 #Atrium_minus2.markers$cluster <- as.character(Atrium_minus2.markers$cluster) 

#Atrium_minus2.markers  <- data.frame(Atrium_minus2.markers[,c(6:7, 5, 2)])
 #Atrium_minus2.markers[,1] <- as.numeric(Atrium_minus2.markers[,1])
#R2<- Atrium_minus2.markers[order(-Atrium_minus2.markers[,1], Atrium_minus2.markers[,3]), ]

markers_Atrium_minus2 <- filter(markers_Atrium_minus, cluster %in%  Atria2)

markers_Atrium_minus2[,1] <- as.numeric(markers_Atrium_minus2[,1])

markers_Atrium_minus2<- markers_Atrium_minus2[order(-markers_Atrium_minus2[,1], markers_Atrium_minus2[,3]), ]
markers_Atrium_minus2 <- subset(markers_Atrium_minus2, p_val_adj<0.05)

markers_Atrium_minus2.to.plot <- unique(markers_Atrium_minus2$gene)

c <- markers_Atrium_minus2.to.plot 
c < factor(c, levels = c)


d2 <- DotPlot(object = Atrium_minus2,  features = rev(c), dot.scale = 4, group.by ='sub_cluster_SAN') + 
  #scale_colour_gradientn(colors = heat_cols) +
  scale_y_discrete(name  = "", 
                   labels = c("Trabecular", "Conduit", "SAN")) +
  theme(legend.position = 'top', text = element_text(size =10), axis.text.x = element_text(size = 10, angle = -45, hjust = 0,vjust = 1), plot.margin = unit(c(0.5,1,0,0), "cm")) +
  xlab('  ')

#d3 <- d2 +  annotate ("rect", xmin = 18.5, xmax = 33.5, ymin = 0, ymax = 3.5, alpha = .1, fill = "blue") 
d4 <- d2 +
  annotate ("rect", xmin = 14.5, xmax = 30.5, ymin = 0, ymax = 3.5, alpha = .1, fill = "blue") 

#d3  / d4 
d4

Idents(Atrium_minus2) <- 'sub_cluster_SAN2'

d_a <- DoHeatmap(object = Atrium_minus2,,  features = rev(c), angle = 90, size = 4)  + NoLegend() 

ggsave(d_a, filename = "/Users/ChristerSylven/Desktop/CMC_ms/aMS/Final/Nov_figs/heatmap_atria_minus.svg",  width = 6, height = 22,  device = "svg", units = "cm")

d_a
```

## Ventricular base, cell cycle phase fractions, mitochondrial and MYH7 and GJA1 variabilities

```{r}
#Percentage of phase fractions
Idents(seu_Ventricle) <- 'ClusterNames'
#Idents(seu_Ventricle) <- 'SCT_snn_res.0.7'
table(Idents(seu_Ventricle))
```

### S_Figure 2B

```{r figs2B, fig.height = 8, fig.width = 12, fig.align = "center"}
seu_Ventricle@meta.data$ClusterNames <- factor(seu_Ventricle@meta.data$ClusterNames, levels = c('Trabecular','Compact',  'Purkinje-related', 'High G2M', 'Exosome-related', 'Outflow tract'))

p10 <- ggplot(seu_Ventricle@meta.data, aes(ClusterNames, fill = base_clusters)) +
  geom_bar(position = "fill") +
  scale_y_continuous(name="Fractions", limits=c(0, 1.0)) +
  theme(text = element_text(size=16),  axis.text.y = element_text(size = 12), axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  xlab(' ') +
  labs(fill = "Base")

p20 <- ggplot(seu_Ventricle@meta.data, aes(ClusterNames, fill = Phase)) +
  geom_bar(position = "fill") +
  scale_y_continuous(name="Fractions", limits=c(0, 1.0)) +
  theme(text = element_text(size=16), axis.text.x = element_text(size = 12, angle = -45, hjust = 0, vjust = 1), axis.text.y = element_text(size = 12)) +
  xlab(' ') +
  labs(fill = "Phase")

p30 <- VlnPlot(seu_Ventricle, features = 'G2M.Score') +
  theme(text = element_text(size=16),  axis.ticks.x = element_blank(),
        axis.text.x = element_blank(), axis.text.y = element_text(size = 12),
        legend.position = "none") +
  xlab(' ')
p40 <- VlnPlot(seu_Ventricle, features = 'S.Score') +
  theme(text = element_text(size=16), axis.text.x = element_text(size = 12, angle = -45, hjust = 0, vjust = 1), axis.text.y = element_text(size = 12), legend.position = "none", plot.margin = unit(c(1, 2, 1, 1), "cm")) +
  xlab(' ')
p50 <- p10/p20
p60 <- p30 / p40
#p5 | p6


# plot only ventricular
#Idents(seu_Muscle) <- 'SCT_snn_res.0.7'
#seu_Ventricle <- subset(seu_Muscle, idents = c(0:2,4,6:7))
#saveRDS(seu_Ventricle, file = 'seu_Ventricle.R')
#seu_Ventricle <- readRDS('seu_Ventricle.R')

# delete clusters w 0 cells
#seu_Ventricle@meta.data[,12] <- factor(seu_Ventricle@meta.data[,12])
#seu_Ventricle@meta.data[,13] <- factor(seu_Ventricle@meta.data[,13])
#seu_Ventricle@meta.data[,15] <- factor(seu_Ventricle@meta.data[,15])


p1 <- ggplot(seu_Ventricle@meta.data, aes(percent.mt, nCount_RNA, colour= base_clusters)) + 
  geom_point(size = 0.75) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") +
  facet_wrap(~ClusterNames, ncol = 3)

p2 <- ggplot(seu_Ventricle@meta.data, aes(percent.mt, G2M.Score, colour= base_clusters)) + 
  geom_point(size = 0.75) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "top") +
  facet_wrap(~ClusterNames, ncol = 3)

p3 <- ggplot(seu_Ventricle@meta.data, aes(percent.mt, S.Score, colour= base_clusters)) + 
  geom_point(size = 0.75) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") +
  facet_wrap(~ClusterNames, ncol = 3)

 p4 <- FeatureScatter(seu_Ventricle, feature1 = "percent.mt", slot = 'data', feature2 = "MYH7", pt.size = 0.75, cols = c('darkmagenta', 'royalblue1', '#FF7F0E', 'blue', 'green',  'magenta'), plot.cor = F) +  theme_gray() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         text = element_text(face = "plain", size = 11),
         axis.text.x=element_text(angle=0, hjust=1, size=10,face="plain"),
         axis.title = element_text(size=10,face="plain"),
         axis.title.y.right = element_text(size = 10,face="plain"),
         legend.text=element_text(size=10),
         legend.title=element_text(size=10),
         legend.position = "none") +
   facet_wrap(~seu_Ventricle@meta.data$ClusterNames, ncol =3) 

 p5 <- FeatureScatter(seu_Ventricle, feature1 = "percent.mt", slot = 'data', feature2 = "GJA1", pt.size = 0.75, cols = c('darkmagenta', 'royalblue1', '#FF7F0E', 'blue', 'green',  'magenta'), plot.cor = F) + 
   theme_gray() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         text = element_text(face = "plain", size = 11),
         axis.text.x=element_text(angle=0, hjust=1, size=11,face="plain"),
         axis.title = element_text(size=11,face="plain"),
         axis.title.y.right = element_text(size = 11,face="plain"),
         legend.text=element_text(size=10),
         legend.title=element_text(size=10),
         legend.position = "none") +
 facet_wrap(~seu_Ventricle@meta.data$ClusterNames, ncol =3) 
   

 layout <- "
ABC
DEF
"

 p1 + p2 + p3 + p4 + p5 + p50 + plot_layout(design = layout)
# p50

#AUC for projection of ventricular clusters on ST map. Tested for 3, 4, 5, 7 and 12top genes where 7 genes gave best resolution
 DE_genes_SC_Muscle_top12 <- extractTopN(seu_Muscle.markers, N = 12)
 
 geneSets <- split(DE_genes_SC_Muscle_top12$gene, paste0('',DE_genes_SC_Muscle_top12$cluster))
 #geneSets
 geneSets <- geneSets[1:8]
 genes <- c("TBX18", "SHOX2", "HCN4") # sinus node
 geneSets <- append(geneSets, list(genes))
 names(geneSets)[9] <- 'SAN'
```


```{r}
# from Asp et al 2019: deposited filtered ST data but not processed
 ST_filtered <- read.table(file = paste0(path1, 
'data2/filtered_matrix.tsv'),  sep = '\t', header = TRUE, check.names=FALSE, stringsAsFactors=FALSE)
 colnames(ST_filtered)[1] <- 'gene_id'
 #dim(ST_filtered)
#Load conversion table used for Asp et al 2019 containing ENSEMBL ids and gene symbols (created from annotation GRCh38_v86 gencode file also used for 10X SC data)
```

, 
```{r}
ensids <- read.table(paste0(path1, "data/genes.tsv"), header = T, stringsAsFactors = F)
#ensids <- read.table("/Users/ChristerSylven/Human-Prenatal-Cardiomyocytes/data/genes.tsv", header = T, stringsAsFactors = F)


 rownames(ensids) <- ensids$gene_id
 
 #Check that all ENSEMBL ids are available in the table

#sum(ST_filtered[,1] %in% ensids$gene_id) == nrow(ST_filtered)
# table(nchar(ensids$gene_id))
 
#merge to get genenames 
ST_filtered <- merge(ST_filtered, ensids, by.x='gene_id', by.y='gene_id')
rownames(ST_filtered) <- ST_filtered[,3114] # genenames
ST_filtered <- ST_filtered[,2:3112]

#Spot names divided into their 3 coordinates
coords <- setNames(data.frame(apply(do.call(rbind, strsplit(colnames(ST_filtered), split = "x")), 2, as.numeric)), nm = c("well", "x", "y"))
#dim(coords)
rownames(coords) <- colnames(ST_filtered) # spot names
#head(coords)

week6_spots <- which(coords$well %in% 5:13) # well 5:13 is week 6 hearts
#length(week6_spots)
ST_6 <- ST_filtered[,week6_spots] #data.frame with week 6 spots

# rownames (genes) used in Asp et al 2019

Article_rownames_ST_6 <- readRDS(paste0(path1, "data/Article_rownames_ST_6.R"))
#length(Article_rownames_ST_6)
ST_6 <- ST_6[Article_rownames_ST_6,] # Genes in ST data used in Asp et al 2019
#dim(ST_6)
cells_rankings_ST_6 <- AUCell_buildRankings(as.matrix(ST_6), nCores = 9, verbose = FALSE)

# save cells_rankings_ST_6 for use in Stereoscope validation
saveRDS(cells_rankings_ST_6, file = paste0(path1, 'cells_rankings_ST_6.R'))
#cells_rankings_ST_6.R <- readRDS(paste0(path1, 'cells_rankings_ST_6.R'))
cells_AUC_ST_6 <- AUCell_calcAUC(geneSets, 
                               cells_rankings_ST_6, 
                               aucMaxRank = ceiling(0.05 * nrow(cells_rankings_ST_6)), verbose = FALSE)
# M for muscle subclusters, SC for base clusters
AUC_ST_m <- setNames(data.frame(t(getAUC(cells_AUC_ST_6))), nm = paste0( 'M', 1:nrow(getAUC(cells_AUC_ST_6))-1))

#on ST tissue 6 weeks
## 6 weeks # data used in Asp et al 2019
g.list <- readRDS(paste0(path1, "data2/grobs.list"))
dims_df <- readRDS(paste0(path1, "data/dims.df"))
```


```{r}
qc_alignment <- read.table(paste0(path1, "data/merged_qc_alignment.txt"))
qc_alignment <- qc_alignment[, c("rep", "x", "y", "new_x", "new_y", "pixel_x", "pixel_y")]

options(repr.plot.width = 13, repr.plot.height = 10) 

#g <- 'MYH6'
#MYH6 <- t(ST_6[g,, drop =F])

coords_6 <- coords[rownames(AUC_ST_m),] 
coords_6 <- cbind(coords_6, AUC_ST_m)
#coords_6 <- cbind(coords_6, MYH6)
#coords_6b <- coords_6[rownames(AUC_ST_m,)]

# save coords_6 for use in Sterescope validtion
saveRDS(coords_6, file = paste0(path1, 'coords_6.R'))

gg_on_tissue <- data.frame(merge( coords_6, qc_alignment, by = "row.names"), row.names = 1)
#gg_on_tissue <- data.frame(merge(gg_on_tissue, AUC_ST_m, by='row.names'))
gg_on_tissue$x <- gg_on_tissue$x.y
gg_on_tissue$y <- gg_on_tissue$y.y
#names(gg_on_tissue)
#fill4=c('darkmagenta', 'royalblue1', '#FF7F0E', 'maroon', 'blue',  'green', 'blue','magenta', 'black', 'white')
colsM <- list()
colsM[[1]] <- c("lightgrey", "lightgrey",  "lightgrey",  "#b55cb5",  'darkmagenta') #trab M0
colsM[[2]] <- c("lightgrey", "lightgrey",  "lightgrey",  "#728ff2",  'royalblue1') #comp M1
colsM[[3]] <- c("lightgrey", "lightgrey",  "lightgrey",  "#ff9a41",  '#FF7F0E') #purkinje M2
colsM[[4]] <- c("lightgrey", "lightgrey",  "lightgrey",  "#fcacac",  'maroon') #atrial trab M3
colsM[[5]] <- c("lightgrey", "lightgrey",  "lightgrey",  "#56baf8",  'blue') #High  G2M M4
colsM[[6]] <- c("lightgrey", "lightgrey",  "lightgrey",  "#ADFF2F",  'green') #atrial cond M5
colsM[[7]] <- c("lightgrey", "lightgrey",  "lightgrey",  "#f0fabe", "lightgreen") #exosomal M6
colsM[[8]] <- c("lightgrey", "lightgrey",  "lightgrey",  "#ffccff",  'magenta') #OFT M7
colsM[[9]] <- c("lightgrey", "lightgrey",  "lightgrey",  "#ff9999",  'red') #SAN M8
colsM[[10]] <- c("lightgrey", "#56baf8",  "#ffccff",  "#ff9999",  'red') #MYH6

names <- c('Trab', 'Comp', 'Purkinje', 'Atrial_trab', 'HighG2M', 'Atrial_cond', 'Exosome', 'OFT', 'SAN') 

#b55cb5 darkmagenta light trab 0
#728ff2 royalblue1 light compact1
#ff9a41 pumpkin #FF7F0E purkinje 2
#fcacac maroon light atrial trab  3
#56baf8 light blue    HighG2M 4
#ADFF2F    green atrial cond 5   
#f0fabe lightgreen light exosomal 6
#ff99ff magenta light      OFT 7
#ff9999 SAN 8


for (j in 1:9) { 
#j=8 # j = 8 is M7 OFT 
  j2 <- j +3

gg_on_tissue$AUC <- gg_on_tissue[,j2]  # 4 - 12 cl 0 - 8  Atrial trabec (M3:7), Atrial conduit (M5:9), SAN (M8:12)  OFT (M7,11) 

gg_on_tissue$AUC2 <- log(gg_on_tissue$AUC+1)
max_val <- max(gg_on_tissue$AUC2)
min_val <- min(gg_on_tissue$AUC2)

#AUC quantitative AUC2 categorical
gg_on_tissue$AUC3 <- cut(gg_on_tissue$AUC2, breaks = c(0, max_val*0.2, max_val*0.4, max_val*0.6, max_val*0.8,  max_val),  labels = c('0', '1', '2', '3', '4'), right = TRUE)
gg_on_tissue$AUC3[is.na(gg_on_tissue$AUC3)] <- 0
#summary(gg_on_tissue$AUC) # linear
#summary(gg_on_tissue$AUC2) # log
#print(j)
#print(summary(gg_on_tissue$AUC3))# categorical based on log


#dev.off()
# projection on slides

cols <- colsM[[j]]
#cols <- colsM[[(9)]]
names(cols) <- paste0(0:4)
#gg_on_tissue$oft <- gg_on_tissue$AUC3
#names(fill4) <- paste0(0:9)
p.list <- lapply(min(unique(gg_on_tissue$rep)):max(unique(gg_on_tissue$well)), function(i) {
  HE_img <- g.list[[i]]
  d <- suppressWarnings(as.numeric(dims_df[i, ]))
  
  g <- rasterGrob(HE_img, width = unit(1, "npc"), height = unit(1, "npc"), interpolate = TRUE)
  
  ggplot(subset(gg_on_tissue, rep == i), aes(pixel_x, d[3] - pixel_y, color = AUC3)) + 
    annotation_custom(g, -Inf, Inf, -Inf, Inf) +
    geom_point(size = 1.5, alpha = 1.0) + # point size 1.5 for small
    # 5.0 for individual
    scale_x_continuous(limits = c(0, d[2]), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, d[3]), expand = c(0, 0)) +
    # categorical values
    scale_color_manual(values = cols) +
    # quantitative values
    #scale_color_gradientn(colours = c("lightgrey", "lightgrey", "lightgrey", "red", "red"), limits = c(min_val, max_val)) +
    theme_void() +
    theme(legend.position="none")
})

#small with 9 sections
p <- cowplot::plot_grid(plotlist = p.list)
#ggsave(p, filename = paste0("/Users/ChristerSylven/Desktop/CMC_ms/GitHub_code/CMC2_git/Figures/High_resolution_figures/S_Figure_9/",names[j], "_v2.pdf"), height = 10, width = 11)
ggsave(p, filename = paste0(path1, "Figures/S_Figure_9/",names[j], "_v2.png"), height = 10, width = 11)
}
```

# Validation of scRNA determined cardiomyocyte clusters by projection of ST maps.

Projections on all nine histologic sections saved on file. Individual sections used in figure 3a also saved on file.

### Figure 3A

Individual section 6 from each ventricular cardiomyocyte type used in figure 1D extracted from saved on file.

```{r}
for (j in 1:9) { 
#j=8 # j = 8 is M7 OFT 
  j2 <- j +3

gg_on_tissue$AUC <- gg_on_tissue[,j2]  # 4 - 12 cl 0 - 8  Atrial trabec (M3:7), Atrial conduit (M5:9), SAN (M8:12)  OFT (M7,11) 

gg_on_tissue$AUC2 <- log(gg_on_tissue$AUC+1)
max_val <- max(gg_on_tissue$AUC2)
min_val <- min(gg_on_tissue$AUC2)

#AUC quantitative AUC2 categorical
gg_on_tissue$AUC3 <- cut(gg_on_tissue$AUC2, breaks = c(0, max_val*0.2, max_val*0.4, max_val*0.6, max_val*0.8,  max_val),  labels = c('0', '1', '2', '3', '4'), right = TRUE)
gg_on_tissue$AUC3[is.na(gg_on_tissue$AUC3)] <- 0
#summary(gg_on_tissue$AUC) # linear
#summary(gg_on_tissue$AUC2) # log
#print(j)
#print(summary(gg_on_tissue$AUC3))# categorical based on log

#dev.off()
# projection on slides

cols <- colsM[[j]]
#cols <- colsM[[(9)]]
names(cols) <- paste0(0:4)
#gg_on_tissue$oft <- gg_on_tissue$AUC3
#names(fill4) <- paste0(0:9)
p.list <- lapply(min(unique(gg_on_tissue$rep)):max(unique(gg_on_tissue$well)), function(i) {
  HE_img <- g.list[[i]]
  d <- suppressWarnings(as.numeric(dims_df[i, ]))
  
  g <- rasterGrob(HE_img, width = unit(1, "npc"), height = unit(1, "npc"), interpolate = TRUE)
  
  ggplot(subset(gg_on_tissue, rep == i), aes(pixel_x, d[3] - pixel_y, color = AUC3)) + 
    annotation_custom(g, -Inf, Inf, -Inf, Inf) +
    geom_point(size = 1.5, alpha = 1.0) + # point size 1.5 for small
    # 5.0 for individual
    scale_x_continuous(limits = c(0, d[2]), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, d[3]), expand = c(0, 0)) +
    # categorical values
    scale_color_manual(values = cols) +
    # quantitative values
    #scale_color_gradientn(colours = c("lightgrey", "lightgrey", "lightgrey", "red", "red"), limits = c(min_val, max_val)) +
    theme_void() +
    theme(legend.position="none")
})

# each section saved alone
#for (i in 1:length(p.list)) {
i = 6 # used for Figure 1d
#pdf(file = paste0("/Users/ChristerSylven/Desktop/CMC_ms/GitHub_code/CMC2_git/Figures/High_resolution_figures/S_Figure_9/individual/",names[j], i, "_v2.pdf"), width = 11, height = 10)
#plot(p.list[[i]])

#png(file = paste0(path1, "Figures/S_Figure_9/individual/",names[j], i, "_v2.png"), width = 11, height = 10)
#plot(p.list[[i]])

#dev.off()

}

#for (j in 1:9) { 
j=8 # j = 8 is M7 OFT 
  j2 <- j +3

gg_on_tissue$AUC <- gg_on_tissue[,j2]  # 4 - 12 cl 0 - 8  Atrial trabec (M3:7), Atrial conduit (M5:9), SAN (M8:12)  OFT (M7,11) 

gg_on_tissue$AUC2 <- log(gg_on_tissue$AUC+1)
max_val <- max(gg_on_tissue$AUC2)
min_val <- min(gg_on_tissue$AUC2)

#AUC quantitative AUC2 categorical
gg_on_tissue$AUC3 <- cut(gg_on_tissue$AUC2, breaks = c(0, max_val*0.2, max_val*0.4, max_val*0.6, max_val*0.8,  max_val),  labels = c('0', '1', '2', '3', '4'), right = TRUE)
gg_on_tissue$AUC3[is.na(gg_on_tissue$AUC3)] <- 0
#summary(gg_on_tissue$AUC) # linear
#summary(gg_on_tissue$AUC2) # log
#print(j)
#print(summary(gg_on_tissue$AUC3))# categorical based on log


#dev.off()
# projection on slides

cols <- colsM[[j]]
#cols <- colsM[[(9)]]
names(cols) <- paste0(0:4)
#gg_on_tissue$oft <- gg_on_tissue$AUC3
#names(fill4) <- paste0(0:9)
p.list <- lapply(min(unique(gg_on_tissue$rep)):max(unique(gg_on_tissue$well)), function(i) {
  HE_img <- g.list[[i]]
  d <- suppressWarnings(as.numeric(dims_df[i, ]))
  
  g <- rasterGrob(HE_img, width = unit(1, "npc"), height = unit(1, "npc"), interpolate = TRUE)
  
  ggplot(subset(gg_on_tissue, rep == i), aes(pixel_x, d[3] - pixel_y, color = AUC3)) + 
    annotation_custom(g, -Inf, Inf, -Inf, Inf) +
    geom_point(size = 3.0, alpha = 1.0) + # point size 1.5 for small
    # 5.0 for individual
    scale_x_continuous(limits = c(0, d[2]), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, d[3]), expand = c(0, 0)) +
    # categorical values
    scale_color_manual(values = cols) +
    # quantitative values
    #scale_color_gradientn(colours = c("lightgrey", "lightgrey", "lightgrey", "red", "red"), limits = c(min_val, max_val)) +
    theme_void() +
    theme(legend.position="none")
})
```

```{r fig3A, fig.height = 5, fig.width = 6, fig.align = "center"}
# each section saved alone
#for (i in 1:length(p.list)) {
i = 3
#pdf(file = paste0("Figures/High_resolution_figures/S_Figure_9/individual/",names[j], i, "_v2.pdf"), width = 11, height = 10)
plot(p.list[[i]])
#dev.off()

#}
```

### Figure 2C

```{r fig2C, fig.height = 5, fig.width = 6, fig.align = "center"}
# Atrial composite atrial trabecular, atrial conduit & SAN
# #plotting atrial spots: trabecular, conduit, both trabecular and conduit, SAN, both conduit & SAN using categorical > 0.6
# with atrial clusters dicotomized
# below is for cl3 (mainly trabecular atrium, val 1 (continous), val 4 categorical( labels 0 & 1)). Then redo for cl5 conduit atrium (val 2 (continous), val 5 categorical (labels 0 & 2)). Val 3 SAN (SHOX2+TBX18+HCN4) related val 3 (continous), val 6 categorical (label 0 & 4). Val7 integrates labels 0 + 1 + 4. 3 identifies spots where both labels 1 & 2 are present

gg_on_tissue$trab <- gg_on_tissue[,7]  
gg_on_tissue$trab <- log(gg_on_tissue$trab+1)
max_val <- max(gg_on_tissue$trab)
min_val <- min(gg_on_tissue$trab)

#AUC quantitative AUC2 categorical
gg_on_tissue$trab <- cut(gg_on_tissue$trab, breaks = c(0, max_val*0.2, max_val*0.4, max_val*0.6, max_val*0.8,  max_val),  labels = c('0', '0', '0', '1', '1'), right = TRUE)
gg_on_tissue$trab[is.na(gg_on_tissue$trab)] <- 0

gg_on_tissue$cond <- gg_on_tissue[,9]  
gg_on_tissue$cond <- log(gg_on_tissue$cond+1)
max_val <- max(gg_on_tissue$cond)
min_val <- min(gg_on_tissue$cond)

#AUC quantitative AUC2 categorical
gg_on_tissue$cond <- cut(gg_on_tissue$cond, breaks = c(0, max_val*0.2, max_val*0.4, max_val*0.6, max_val*0.8,  max_val),  labels = c('0', '0', '0', '2', '2'), right = TRUE)
gg_on_tissue$cond[is.na(gg_on_tissue$cond)] <- 0

gg_on_tissue$SAN <- gg_on_tissue[,12]  
gg_on_tissue$SAN <- log(gg_on_tissue$SAN+1)
max_val <- max(gg_on_tissue$SAN)
min_val <- min(gg_on_tissue$SAN)

#AUC quantitative AUC2 categorical
gg_on_tissue$SAN <- cut(gg_on_tissue$SAN, breaks = c(0, max_val*0.2, max_val*0.4, max_val*0.6, max_val*0.8,  max_val),  labels = c('0', '0', '0', '3', '3'), right = TRUE)
gg_on_tissue$SAN[is.na(gg_on_tissue$SAN)] <- 0
#table(gg_on_tissue$trab)
#table(gg_on_tissue$cond)
#table(gg_on_tissue$SAN)

gg_on_tissue$atria <- as.character(gg_on_tissue$trab)
gg_on_tissue$atria[which(gg_on_tissue$cond =='2')] <- '2'
gg_on_tissue$atria[which(gg_on_tissue$cond =='2' & gg_on_tissue$trab == '1')] <- '3' 
gg_on_tissue$atria[which(gg_on_tissue$SAN =='3')] <- '5'
gg_on_tissue$atria[which(gg_on_tissue$cond =='2' & gg_on_tissue$SAN== '3')] <- '4'
gg_on_tissue$atria[which(gg_on_tissue$trab =='1' & gg_on_tissue$SAN== '3')] <- '6'
gg_on_tissue$atria[which(gg_on_tissue$trab =='1' & gg_on_tissue$cond =='2' & gg_on_tissue$SAN== '3')] <- '7'
#table(gg_on_tissue$atria)
gg_on_tissue$atria <- as.factor(gg_on_tissue$atria)


# list all maps categorical color
cols <- c("lightgrey", "blue", "green", "orange", 'red', "red", "red", 'red')
names(cols) <- paste0(0:7)

p.list <- lapply(min(unique(gg_on_tissue$rep)):max(unique(gg_on_tissue$rep)), function(i) {
  HE_img <- g.list[[i]]
  d <- suppressWarnings(as.numeric(dims_df[i, ]))
  
  g <- rasterGrob(HE_img, width = unit(1, "npc"), height = unit(1, "npc"), interpolate = TRUE)
  
  ggplot(subset(gg_on_tissue, rep == i), aes(pixel_x, d[3] - pixel_y, color = atria)) + 
    annotation_custom(g, -Inf, Inf, -Inf, Inf) +
    geom_point(size = 3.0, alpha = 1) + #size = 1.5 for small, 3.0 for individual
    scale_x_continuous(limits = c(0, d[2]), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, d[3]), expand = c(0, 0)) +
    scale_color_manual(values = cols) +
    theme_void() 
})
# all 9 sections use point size = 1.5
#p <- cowplot::plot_grid(plotlist = p.list)
#ggsave("/Users/ChristerSylven/Desktop/CMC_ms/Figures/High_resolution_figures/Figure_2/atria_small_v2.pdf", p, height = 18, width = 22,  device = "pdf", dpi = 'retina', units = "cm")
#dev.off()

#only section 9 where SAN is use point size = 3.0
plot(p.list[[9]])
#ggsave("/Users/ChristerSylven/Desktop/CMC_ms/Figures/High_resolution_figures/Figure_2/aFigure2c_v2.pdf", e, width = 22, height = 18,  device = "pdf", dpi = 'retina', units = "cm")
#dev.off()


```

### Manual identification of spots with OFT cardiomyocytes (3 + 4) in section 7.

In order to select spots in outflow tract that contain OFT CMC for analysis of colocalized non CM cells in section 7.

```{r fig4A, fig.height = 5, fig.width = 5, fig.align = "center"}
# proportions of base non cardiomyocyte clusters located in OFT and atrial  conduit and SAN-enriched spots.
# First locate manually OFT, atrial conduit and SAN spots in section 7 (OFT) and section 13 (Atrial)
#table(coords_6$well)
# 5   6   7   8   9  10  11  12  13 
#103 104 159 186 214 209 178 182 180 
#cols for OFT:
cols <- c("lightgrey", "lightgrey", "lightgrey",   "#ffccff",   "magenta")
names(cols) <- paste(0:4)
#cols for Atria:
cols <- c("lightgrey", "blue", "green", "orange", 'red', "red", "red", 'red')
names(cols) <- paste0(0:7)

# to identify spots with OFT (section 3, well/rep 7), atrial conduit section 9 (well/rep13)
# OFT generated as AUC3 in loop
gg_on_tissue$oft <- gg_on_tissue$AUC3
#names(gg_on_tissue)

f <- ggplot(
  subset(gg_on_tissue, rep == 7), aes(x, 36 - y, color = oft)) +
  geom_point(size = 1.5) +
  scale_x_continuous(limits = c(1, 33)) +
  scale_y_continuous(limits = c(1, 35)) + 
  theme_void() +
  theme(legend.position="none")+
  scale_color_manual(values = cols) +
  theme_void() 
#f
ggplotly(f)

#OFT section 7 spots:
OFT_spots_section3_edited <- c('7x15x20', '7x15x21', '7x16x18','7x16x19', '7x16x20','7x16x21', '7x16x23','7x16x24',  '7x17x19','7x17x20', '7x17x21', '7x17x22', '7x17x23', '7x17x24','7x18x19', '7x18x21','7x18x19','7x18x22','7x18x23', '7x18x24', '7x18x25', '7x19x20', '7x19x22', '7x19x23', '7x19x24', '7x19x25', '7x20x21', '7x20x22', '7x20x23', '7x20x24', '7x20x25','7x21x21', '7x21x23', '7x21x24', '7x22x23', '7x22x24')

length(OFT_spots_section3_edited)
#Atria only conduit section 13 spots:
#Conduit_spots_section9_edited <- c('13x23x9','13x22x10',  '13x21x11', '13x21x13', '13x20x16', '13x20x11', '13x19x12', '13x18x13','13x18x17','13x18x18', '13x16x16')

#markers: for base sc clusters
#seu.markers <- FindAllMarkers(seu, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
#DE_genes_SC_top12 <- extractTopN(seu.markers, N = 12)
#DE_genes_SC_top12$cluster <- as.character(DE_genes_SC_top12$cluster) 
#DE_genes_SC_top12 <- data.frame(DE_genes_SC_top12[,c(6:7, 5, 2)])
```

# Proportions of non CM cell types in OFT CM spots

### Figure 3B

```{r}
# #markers: for base sc clusters to use for corrplot
Idents(seu) <- 'SCT_snn_res.0.38'
seu.markers <- FindAllMarkers(seu, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
DE_genes_SC_top12 <- extractTopN(seu.markers, N = 12)
DE_genes_SC_top12$cluster <- as.character(DE_genes_SC_top12$cluster) 
#DE_genes_SC_top12_all_17_clusters <- DE_genes_SC_top12
#geneSets_SC_all_17_clusters <- split(DE_genes_SC_top12_all_17_clusters$gene,DE_genes_SC_top12_all_17_clusters$cluster)
#geneSets_SC_all_17_clusters <- geneSets_SC_all_17_clusters[c(1:2,10:17,3:9)]
#saveRDS(geneSets_SC_all_17_clusters, file = '/Users/ChristerSylven/Desktop/CMC_ms/GitHub_code/data/geneSets_SC_all_17_clusters.R')

#DE_genes_SC_top12 <- readRDS('/Users/ChristerSylven/Desktop/CMC_ms/GitHub_code/data/geneSets_SC_all_17_clusters.R')
# find categorical AUC for the 14 non cardiomyocyte scRNA clusters given in markers2
DE_genes_SC_top12 <- subset(DE_genes_SC_top12, !cluster ==1)
DE_genes_SC_top12 <- subset(DE_genes_SC_top12, !cluster == 8)
DE_genes_SC_top12 <- subset(DE_genes_SC_top12, !cluster ==13)

DE_genes_SC_top12[,1] <- factor(DE_genes_SC_top12[,1], levels=unique(DE_genes_SC_top12[,1])) # delete rows w zero cells
geneSets_SC <- split(DE_genes_SC_top12$gene,DE_genes_SC_top12$cluster)

geneSets_SC <- geneSets_SC[c(1,8:14,2:7)]
#geneSets_SC

cells_AUC_ST_6_SC <- AUCell_calcAUC(geneSets_SC, 
                                 cells_rankings_ST_6, 
                                 aucMaxRank = ceiling(0.05 * nrow(cells_rankings_ST_6)), verbose = FALSE)

AUC_ST_SC <- setNames(data.frame(t(getAUC(cells_AUC_ST_6_SC))), nm = paste0( 1:nrow(getAUC(cells_AUC_ST_6_SC))-1))
colnames(AUC_ST_SC) <- c('0', '2', '3', '4', '5', '6', '7', '9', '10', '11', '12', '14', '15', '16')

# add column with section identity = rep
#AUC_ST_cat_a <- cbind(AUC_ST_cat, gg_on_tissue[,17])
#colnames(AUC_ST_cat_a)[15] <- 'rep'
# to select only slide 13 where SAN is expressed
#AUC_ST_cat_b <- subset(AUC_ST_cat_a,rep == '13')
#AUC_ST_cat <- AUC_ST_cat_b[,1:14]

AUC_ST_SC_cat <- log(AUC_ST_SC+1) # 
min_val <- apply(AUC_ST_SC_cat, 2, min)
max_val<- apply(AUC_ST_SC_cat, 2, max)

#for ( i in 1:21) {AUC_ST_cat[,i] = AUC_ST_cat[,i]/(max_val[i]-min_val[i])}
for ( i in 1:14) {
  max_val<- max(AUC_ST_SC_cat[,i])
  AUC_ST_SC_cat[,i]<- cut(AUC_ST_SC_cat[,i], breaks = c(0, max_val*0.2, max_val*0.4, max_val*0.6,   max_val*0.8, max_val), labels = c('0', '1', '2', '3', '4'), right = TRUE)
}
AUC_ST_SC_cat[is.na(AUC_ST_SC_cat)] <- 0 

OFT <- as.data.table(AUC_ST_SC_cat[OFT_spots_section3_edited, ])
#summary(OFT) 
#nrow(OFT)

test <- as.numeric(as.data.frame(tstrsplit(summary(OFT)[1,], ":", fixed=TRUE))[,2])
test <- as.data.frame(cbind(c(0,2,3,4,5,6,7,9,10,11,12,14,15,16), test))
rownames(test) <- test[,1] 
test$firstq <-nrow(OFT) -test[,2]
test <- test[,-1]
#test
test1_4s <- c((test[1,2]+test[8,2])/2, test[2,2], test[3,2], test[4,2], (test[5,2]+ test[10,2])/2, test[6,2], (test[7,2]+test[9,2])/2 , test[11,2], test[12,2], test[13,2], test[14,2]) 
OFT1_4_s <- test1_4s 

OFT1_4_s <- OFT1_4_s/nrow(OFT)

labels_s2 = c("Endothel", # E
              "Epicardium_d",# I
              "Fb_skeleton",#A
              "Fb_s_vascular",# G
              "Erythrocyte",# C
              "Fb_l_vascular",# B
              "Fb_smooth_m",# F
              "Epicardium",# D
              "Immune",# H
              "Schwann",# J
              "Neural crest" # K
)
# colors for the merged cell types (rev order for plot
                           
fill_ss <- c(
'green',  # epic
'#9EDAE5', # Cardiac neural crest
'#BCBD22', # Immune
'red',  # erythrc
'#2CA02C',  # smaller vasc
'#17BECF', # Schwann
'#FFBB78', # card skele 
'#1F77B4', # endoth
'#F7B6D2', # # Fb_l_vasc
'#8C564B', # smooth musc 
'#FF7F0E' # EPDC
) 

# non-CM in OFT1_4__s spots,
n_s <- data.frame(labels_s2, OFT1_4_s)
colnames(n_s) <- c('cluster', 'fraction')

n_s <- n_s[order(-n_s[,2]),]
n_s[,1] <- as.factor(n_s[,1])
n_s$cluster= with(n_s, reorder(cluster, fraction))
ggplot(n_s, aes(x=cluster, y= fraction, fill = cluster)) + 
  geom_col() +
  scale_fill_manual(values=fill_ss) +
  #scale_fill_manual(values=c('black',"#009E73", "#D55E00", "#F0E442", "greenyellow",'red',"orange", "gray50", "#009E73", "#F0E442", "orange")) +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 1, size = 16), axis.text.y =   element_text(size   = 16), axis.title=element_text(size=16,face="bold"), legend.position = 'none', legend.title = element_text( size = 16), plot.margin = unit(c(1, 6, 1, 1), "cm")) +
  ylim(0,1.0) +
  xlab(NULL)+
  coord_flip()
```

# Proportions of non CM cell types in atrial Conduit CM spots

```{r}


Conduit2 <- subset(gg_on_tissue, atria == '2')
Conduit <- as.data.table(AUC_ST_SC_cat[rownames(Conduit2), ])
#summary(Conduit) 
nrow(Conduit) #46
```

### Figure 4A

```{r}

test <- as.numeric(as.data.frame(tstrsplit(summary(Conduit)[1,], ":", fixed=TRUE))[,2])
test <- as.data.frame(cbind(c(0,2,3,4,5,6,7,9,10,11,12,14,15,16), test))
rownames(test) <- test[,1] 
test$firstq <-nrow(Conduit) -test[,2]
test <- test[,-1]
#test 1+8, 5+10 7+9, 
test1_4s <- c((test[1,2]+test[8,2])/2, test[2,2], test[3,2], test[4,2], (test[5,2]+ test[10,2])/2, test[6,2], (test[7,2]+test[9,2])/2 , test[11,2], test[12,2], test[13,2], test[14,2]) 
Conduit1_4_s <- test1_4s 

Conduit1_4_s <- Conduit1_4_s/nrow(Conduit)

labels_s2 = c("Endothel", # E
              "Epicardium_d",# I
              "Fb_skeleton",#A
              "Fb_s_vascular",# G
              "Erythrocyte",# C
              "Fb_l_vascular",# B
              "Fb_smooth_m",# F
              "Epicardium",# D
              "Immune",# H
              "Schwann",# J
              "Neural crest" # K
)
# colors for the merged cell types (rev order for plot

fill_s_c <- c(
  '#BCBD22', # Immune
  '#9EDAE5', # Cardiac neural crest
  '#2CA02C',  # smaller vasc
  '#FFBB78', # card skele 
  '#17BECF', # Schwann
  'green',  # epic
  '#8C564B', # smooth musc 
  'red',  # erythrc
  '#F7B6D2',  # larger vasc
  '#FF7F0E', # EPDC
  '#1F77B4') # endoth

# non-CM in  Cond_s & SAN_s spots
n_s <- data.frame(labels_s2, Conduit1_4_s)
colnames(n_s) <- c('cluster', 'fraction')


n_s <- n_s[order(-n_s[,2]),]
n_s[,1] <- as.factor(n_s[,1])
n_s$cluster= with(n_s, reorder(cluster, fraction))
ggplot(n_s, aes(x=cluster, y= fraction, fill = cluster)) + 
  geom_col() +
  scale_fill_manual(values=fill_s_c) +
  #scale_fill_manual(values=c('black',"#009E73", "#D55E00", "#F0E442", "greenyellow",'red',"orange", "gray50", "#009E73", "#F0E442", "orange")) +
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 1, size = 16), axis.text.y =   element_text(size   = 16), axis.title=element_text(size=16,face="bold"), legend.position = 'none', legend.title = element_text( size = 16), plot.margin = unit(c(1, 6, 1, 1), "cm")) +
  ylim(0,1.0) +
  xlab(NULL)+
  coord_flip()

#Trabecular2 <- subset(gg_on_tissue, atria == '1')

#Trabecular <- as.data.table(AUC_ST_SC_cat[rownames(Trabecular2), ])
#summary(Trabecular) 
#nrow(Trabecular) #46

#test <- as.numeric(as.data.frame(tstrsplit(summary(Trabecular)[1,], ":", fixed=TRUE))[,2])
#test <- as.data.frame(cbind(c(0,2,3,4,5,6,7,9,10,11,12,14,15,16), test))
#rownames(test) <- test[,1] 
#test$firstq <-nrow(Trabecular) -test[,2]
#test <- test[,-1]
#test 1+8, 5+10 7+9, 
#test1_4s <- c((test[1,2]+test[8,2])/2, test[2,2], test[3,2], test[4,2], (test[5,2]+ test[10,2])/2, test[6,2], (test[7,2]+test[9,2])/2 , test[11,2], test[12,2], test[13,2], test[14,2]) 
#Trabecular1_4_s <- test1_4s 

#Trabecular1_4_s <- Trabecular1_4_s/nrow(Trabecular)

labels_s2 = c("Endothel", # E
              "Epicardium_d",# I
              "Fb_skeleton",#A
              "Fb_s_vascular",# G
              "Erythrocyte",# C
              "Fb_l_vascular",# B
              "Fb_smooth_m",# F
              "Epicardium",# D
              "Immune",# H
              "Schwann",# J
              "Neural crest" # K
)
# colors for the merged cell types (rev order for plot

fill_s_c <- c(
  '#FFBB78', # card skele
  '#9EDAE5', # Cardiac neural crest
  '#17BECF', # Schwann
  '#BCBD22', # Immune
  '#2CA02C',  # smaller vasc
  'green',  # epic
  '#8C564B', # smooth musc 
  '#F7B6D2',  # larger vasc
  '#FF7F0E', # EPDC
  'red',  # erythrc
  '#1F77B4') # endoth
# colors for the merged cell types

# non-CM in OFT1_4__s spots, Cond_s & SAN_s
#n_s <- data.frame(labels_s2, Trabecular1_4_s)
#colnames(n_s) <- c('cluster', 'fraction')

#n_s <- n_s[order(-n_s[,2]),]
#n_s[,1] <- as.factor(n_s[,1])
#n_s$cluster= with(n_s, reorder(cluster, fraction))
#ggplot(n_s, aes(x=cluster, y= fraction, fill = cluster)) + 
#  geom_col() +
#  scale_fill_manual(values=fill_s_c) +
  #scale_fill_manual(values=c('black',"#009E73", "#D55E00", "#F0E442", "greenyellow",'red',"orange", "gray50", "#009E73", "#F0E442", "orange")) +
#  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = 1, size = 16), axis.text.y =   element_text(size   = 16), axis.title=element_text(size=16,face="bold"), legend.position = 'none', legend.title = element_text( size = 16), plot.margin = unit(c(1, 6, 1, 1), "cm")) +
#  ylim(0,1.0) +
#  xlab(NULL)+
#  coord_flip()

#statistics
f <- GetAssayData(object = seu_Ventricle, slot = 'counts')
f2 <- NormalizeData(f)
test = 'GJA1'

seu_Ventricle$test  <- f2[test,]
#class(seu_Ventricle$test)
#head(seu_Ventricle$test)
#summary(seu_Ventricle$test)
#seu_Ventricle$test


# extract markers for xls file & GO
Idents(seu_Muscle) <- 'sub_cluster_SAN'
p.val = p.val.threshold = 0.01

seu_Muscle.markers <- seu_Muscle.markers %>% 
  group_by(cluster) %>%
  arrange(-avg_log2FC) %>%
  mutate(or = 1:n()) %>%
  mutate(sign = ifelse(or %in% 1:10, gene, ""), DE = ifelse(p_val_adj < p.val.threshold, paste0("adj. P-val < ", p.val.threshold), paste0("adj. P-val >= ", p.val.threshold)),
         sign.flat = ifelse(or %in% (n() - 10):n() | or %in% 1:10, gene, "")) %>%
  arrange(cluster)
#head(seu_Muscle.markers)

wb <- createWorkbook()
 
seu_Muscle <- readRDS(paste0(path1, 'seu_Muscle_2021_May.R'))
Idents(seu_Muscle) <- 'sub_cluster_SAN'

n <- nlevels(Idents(seu_Muscle))-1

#n 

for(i in 0:n){
name <- paste0('l',i)

#print(paste0('Cluster ', i))

t <- FindMarkers(object = seu_Muscle, ident.1 = i, min.pct = 0.1)
t <- t %>% filter(p_val_adj <0.01)
t$gene <- rownames(t)
t <- bitr(t$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")
t <- list(t[,2])
assign(name, t, envir = globalenv())
}

# definition of SAN according to literature. Conservative, but may be used as there are too few SAN cells. References in article 
eg9 = bitr( c('TBX18', 'SHOX2', 'HCN4'), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")
l9 = list(eg9[,2])

list_SC_Ventricle <- c(l0, l1, l2, l4, l6, l7)
names(list_SC_Ventricle) <- c("Trabecular", "Compact",  "Purkinje-related",   "High G2M", "Exosome-related",  "Outflow tract")
# to keep input order in output
names(list_SC_Ventricle) <- factor(c("Trabecular", "Compact",  "Purkinje-related",   "High G2M", "Exosome-related",  "Outflow tract"), levels = c("Trabecular", "Compact",  "Purkinje-related",   "High G2M", "Exosome-related",  "Outflow tract")) 

```

# GO terms

Choose which ontology BP = biological processes, MF = molecular function, CC = cellular component

## Edited GO BP terms for ventricular clusters

```{r}
ck <- compareCluster(geneCluster = list_SC_Ventricle, fun = "enrichGO", OrgDb = org.Hs.eg.db, ont = "BP", pvalueCutoff = 0.05)
#ck <- compareCluster(geneCluster = list_SC_Muscle, fun = "enrichGO", OrgDb = org.Hs.eg.db, ont = "MF", pvalueCutoff = 0.05)
#ck <- compareCluster(geneCluster = list_SC_Muscle, fun = "enrichGO", OrgDb = org.Hs.eg.db, ont = "CC", pvalueCutoff = 0.05)

#head(as.data.frame(ck))

#fig.width=6, fig.height=6, fig.retina=2}
#knitr::opts_chunk$get('fig.retina')
# plot

#dotplot(ck, font.size = 12) 
#+
 # theme(axis.text.x = element_text(size = 11, angle = 90, vjust = 1, hjust = 1), axis.text.y = element_text(size = 13, angle = 0, hjust = 1,  colour = 'black')) +
#  ggtitle("")

```


### Figure 1C

```{r fig1C, fig.height = 8, fig.width = 6, fig.align = "left"}
#For editing. Print descriptions for respective cluster and select relevant descriptions and their order in dotplot.
#head(as.data.frame(ck))
ck2 <- as.data.frame(ck)

#print Descriptions for cluster in order to select relevant GO terms
ck3 <- filter(ck2, Cluster == 'Outflow tract')
ck4 <- ck3[ c('Description','p.adjust')]
#ck4[1:100,]

ventricle <- c("cardiac muscle tissue development", 
               "muscle organ development",
           "heart contraction",
           "cardiac muscle tissue morphogenesis", 
           "regulation of muscle contraction", 
           "regulation of cardiac conduction",
           "second-messenger-mediated signaling",
           "parasympathetic nervous system development",
           "regulation of calcium-mediated signaling",
           "cell communication involved in cardiac conduction",
           "regulation of membrane potential",
           "regulation of transmembrane transporter activity",
           "regulation of heart rate",
           "glycolytic process", 
           "glucose catabolic process to pyruvate",
           "NADH regeneration",
           "response to hypoxia",
           "cotranslational protein targeting to membrane", 
           "protein targeting to ER",
           "extracellular structure organization",
           "regulation of actin cytoskeleton organization", 
           'mesenchyme development',
           'aminoglycan catabolic process',
           "regulation of mitochondrion organization",  
          "nuclear division",
          "organelle fission",
          "chromosome segregation",
          "mitotic spindle organization",                                     
          "cell cycle G2/M phase transition",
          "regulation of mitotic cell cycle phase transition",
          "G2/M transition of mitotic cell cycle")
# plot ventricular clusters
ck2 <- as.data.frame(ck)
# transform GeneRatio (character, i.e. 6/98) to numeric
z <- apply(do.call(rbind, strsplit(ck2$GeneRatio, split = "/")), 2, as.numeric)
colnames(z) <- c( "x", "y")
z <- as.data.frame(z)
#dim(z)
#head(z)
ck2$GeneRatio <- z$x/z$y

ck3 <- filter(ck2, Description %in% ventricle)

#ck3 <- ck2[ck2$Description %in% devA,]
#dim(ck3)

ck3$Description <- factor(ck3$Description, levels = ventricle)

labels_ventricular <- c(
  c("Cardiac muscle tissue development",
    "Muscle organ development",
    "Heart contraction",
    "Cardiac muscle tissue morphogenesis", 
    "Regulation of muscle contraction", 
    "Regulation of cardiac conduction",
    "Second-messenger-mediated signaling",
    "Parasympathetic nervous system development",
    "Cell communication involved in cardiac conduction",
    "Regulation of membrane potential",
    "Regulation of transmembrane transporter activity",
    "Regulation of heart rate",
    "Glycolytic process", 
    "Glucose catabolic process to pyruvate",
    "NADH regeneration",
    "Response to hypoxia",
    "Cotranslational protein targeting to membrane", 
    "Protein targeting to ER",
    "Extracellular structure organization",
    "Regulation of actin cytoskeleton organization",
    'Mesenchyme develoment',
    'Aminoglycan catabolic process',
    "Nuclear division",
    "Organelle fission",
    "Chromosome segregation",
    "Mitotic spindle organization",                                           
    "Cell cycle G2/M phase transition",
    "Regulation of mitotic cell cycle phase transition",
    "G2/M transition of mitotic cell cycle")
)

ggplot(ck3, aes(x = Cluster, y = Description)) + 
  #p <- ggplot(ck3, aes(x = Cluster, y = reorder(Description, desc(Description)))) + 
  geom_point(aes(size = GeneRatio, color = p.adjust)) +
  theme_bw(base_size = 16) +
  scale_colour_gradient(limits = c(0, 0.05), low = "red") +
  guides(
    #reverse color order (higher value on top)
    color = guide_colorbar(reverse = TRUE),
    #reverse size order (higher diameter on top) 
    size = guide_legend(reverse = TRUE))+
  ylab(NULL) +
  xlab(NULL) +
  scale_y_discrete(labels = labels_ventricular) +
  theme_gray() +
  theme(axis.text.x = element_text(size = 13, angle = 90, vjust = 1, hjust = 1), axis.text.y = element_text(size = 11, angle = 0, hjust = 1, vjust = 1, colour = 'black')) +
  ggtitle("")
 # theme(legend.title = element_text(size=12, vjust = -25, hjust = -25)) 
#+
 # theme(plot.margin=unit(c(1,1,1,1),"cm"))

```

## Edited GO BP terms for atrial clusters

### Figure 2B

```{r fig2B, fig.height = 6, fig.width = 6, fig.align = "left"}


list_SC_Atria <- c(l3, l5, l9)
names(list_SC_Atria) <- c("Atrial trabecular",  "Atrial conduit", "SAN")

ck <- compareCluster(geneCluster = list_SC_Atria, fun = "enrichGO", OrgDb = org.Hs.eg.db, ont = "BP", pvalueCutoff = 0.05)
#ck <- compareCluster(geneCluster = list_SC_Muscle, fun = "enrichGO", OrgDb = org.Hs.eg.db, ont = "MF", pvalueCutoff = 0.05)
#ck <- compareCluster(geneCluster = list_SC_Muscle, fun = "enrichGO", OrgDb = org.Hs.eg.db, ont = "CC", pvalueCutoff = 0.05)

#head(as.data.frame(ck))

# plot

#dotplot(ck, font.size = 12) 
#+
#  theme(axis.text.x = element_text(size = 11, angle = 90, vjust = 1, hjust = 1), axis.text.y = element_text(size = 13, angle = 0, hjust = 1,  colour = 'black')) +
#  ggtitle("")

#print Descriptions for cluster in order to select relevant GO terms
#head(as.data.frame(ck))
ck2 <- as.data.frame(ck)

#print Descriptions for cluster in order to select relevant GO termsck3 <- filter(ck2, Cluster == 'Outflow tract')
ck4 <- ck3[ c('Description','p.adjust')]
#ck4[1:100,]
ck3 <- filter(ck2, Cluster == 'Atrial conduit')
ck4 <- ck3[ c('Description','p.adjust')]
#ck4[1:100,]

atria <- c(
  'SA node cell to atrial cardiac muscle cell communication',
  'cardiac conduction system development',
  'regulation of action potential',
  'regulation of heart rate',
  'cell-cell signaling involved in cardiac conduction',
  'cell communication by electrical coupling',
  'muscle contraction',
  'muscle system process',
  'cardiac septum morphogenesis',
  'tissue migration',
  'epithelial cell migration',
  'positive regulation of epithelial to mesenchymal transition',
  'mesenchymal cell development',
  'endothelial cell proliferation', 
  'connective tissue development',
  'BMP signaling pathway',
  'response to transforming growth factor beta',
  'regulation of canonical Wnt signaling pathway')

ck2 <- as.data.frame(ck)
# transform GeneRatio (character, i.e. 6/98) to numeric
z <- apply(do.call(rbind, strsplit(ck2$GeneRatio, split = "/")), 2, as.numeric)
colnames(z) <- c( "x", "y")
z <- as.data.frame(z)

ck2$GeneRatio <- z$x/z$y

ck3 <- filter(ck2, Description %in% atria)

ck3$Description <- factor(ck3$Description, levels = atria)

# plot atrial clusters
labels_atrial <-
  c(
    'SA node cell to atrial cardiac muscle cell communication',
    'Cardiac conduction system development',
    'Regulation of action potential',
    'Regulation of heart rate',
    'Cell-cell signaling involved in cardiac conduction',
    'Cell communication by electrical coupling',
    'Muscle contraction',
    'Muscle system process',
    'Cardiac septum morphogenesis',
    'Tissue migration',
    'Epithelial cell migration',
    'Positive regulation of epithelial to mesenchymal transition',
    'Mesenchymal cell development',
    'Endothelial cell migration',
    'Connective tissue development',
    'BMP signaling pathway',
    'Response to transforming growth factor beta',
    'Regulation of canonical Wnt signaling pathway')

ggplot(ck3, aes(x = Cluster, y = Description)) + 
  #p <- ggplot(ck3, aes(x = Cluster, y = reorder(Description, desc(Description)))) + 
  geom_point(aes(size = GeneRatio, color = p.adjust)) +
  theme_bw(base_size = 16) +
  scale_colour_gradient(limits = c(0, 0.05), low = "red") +
  guides(
    #reverse color order (higher value on top)
    color = guide_colorbar(reverse = TRUE),
    #reverse size order (higher diameter on top) 
    size = guide_legend(reverse = TRUE))+
  ylab(NULL) +
  xlab(NULL) +
  scale_y_discrete(labels = labels_atrial) +
  theme_gray() +
  theme(axis.text.x = element_text(size = 11, angle = 90, vjust = 1, hjust = 1), axis.text.y = element_text(size = 11, angle = 0, hjust = 1,  colour = 'black')) +
 #scale_y_discrete(position = "right") +
  ggtitle("")
#+
 # theme(plot.margin=unit(c(1,1,1,1),"cm"))
      
#   theme(legend=element_text(size=15,  vjust=-25, hjust = 1.1), legend.box = 'horizontal') +
#  theme(plot.margin = unit(c(5,1,1,1), "cm"))
 
```

# Validation of AUC with Stereoscope

### Figure S 5

```{r figS5, fig.height = 8, fig.width = 8, fig.align = "center"}
# Collect tsv files with cell type proportions
tsv.files <- list.files(paste0(path1, 'fh_Christer/'), pattern = ".tsv", recursive = TRUE, full.names = TRUE)

# with rownames (spot names as in original ST map)#find subset stereoscope section spots  w unique correspondence to article ST spots spots stereoscope 1-8. 1 = ST 13
# 1 -> 13
# 2 -> 8
# 3 -> 11
# 4 -> 7
# 5 -> 12
# 6 -> 9
# 7 -> 6 
# 8 -> 10

scMat_5 <- do.call(rbind, lapply(seq_along(tsv.files), function(i) {
  # read one matrix at the time
  df <- read.table(file = tsv.files[i])
  if(i ==1) {
    j = '13'
  } 
  if (i == 2) {
    j = 8
 }
  if (i == 3) {
    j = 11
  }
   if (i == 4) {
     j = 7
 }
   if (i == 5) {
     j = 12
 }
   if (i == 6) {
     j = 9
 }
   if (i == 7) {
     j = 6
  }
   if (i == 8) {
     j = 10
 }
  # add a sample number to the rownames
 rownames(df) <- paste(j, gsub(pattern = "X", replacement = "", x = rownames(df)), sep = "x")
  return(df)
}))
# order of clusters as in SC DimPlot
scMat_5 <- scMat_5[,c(2,15,9, 6, 11, 14, 8, 1, 10, 5, 4, 7, 13, 12, 3 )]

#dim(scMat_5) #1375 15

#coords_scMat_5 <- setNames(data.frame(apply(do.call(rbind, strsplit(rownames(scMat_5), split = "x")), 2, as.numeric)), nm = c("rep", "x", "y"))
#rownames(coords_scMat_5) <- rownames(scMat_5)
 
coords_6 <- readRDS(paste0(path1,'coords_6.R'))
#dim(coords_6) #1515

scMat_5 <- subset(scMat_5, rownames(scMat_5) %in% rownames(coords_6))
#dim(scMat_5) #1357

# use same 15 scRNA clusters as used in steroscope to generate AUC profiles

# old Seurat analysis, 14 SC clusters
#seu2 <- readRDS('/Users/ChristerSylven/Desktop/aLudvig_jan11/SC_data_processing_and_analysis/results/matrices/Seurat_SC_doublet_filtered')
#seu2 <- UpdateSeuratObject(seu2)
#seu2
#names(seu2)
#Idents(seu2) <- 'res.0.7'
#DimPlot(seu2, reduction = 'tsne', label = TRUE)
#names(seu2[[]])
#table(seu2@meta.data[,8])


#DE genes 15 clusters for old Seurat
#DE_genes_SC_old <- data.frame() # Create empty data.frame
#for (i in 1:15) { # Iterate through the numbers of clusters
  # 10 clusters 
#  excel_sheet <- read.xlsx("/Users/ChristerSylven/Desktop/aLudvig_jan11/SC_data_processing_and_analysis/results/matrices/marker_genes_from_Seurat_2.xlsx", sheet = i) # Read sheet i
#  DE_genes_SC_old <- rbind(DE_genes_SC_old, excel_sheet) # Bind the new data.frame called excel_sheet to the data.frame called DE_genes_SC
#}
#dim(DE_genes_SC_old)
#names(DE_genes_SC_old)
#extractTopN_old <- function(DE_genes, N = 5) { # use avg_log2FC for new & avg_logFC for old Seurat
#  subset(DE_genes, avg_logFC > 0) %>% # Subset the data.frame to include only positive avg_logFC values
#    arrange(-avg_logFC) %>% # Arrange the data.frame by decreasing avg_logFC ("-" = decreasing)
#    group_by(cluster) %>% # Group the data.frame by cluster (this will produce a "grouped" data.frame, i.e. a subset for each cluster)
#    top_n(N, avg_logFC) %>% # Extract top N rows for each cluster based on avg_logFC
#    arrange(cluster) 
#}

extractTopN <- function(DE_genes, N = 5) { # use avg_log2FC for new & avg_logFC for old Seurat
  subset(DE_genes, avg_log2FC > 0) %>% # Subset the data.frame to include only positive avg_logFC values
    arrange(-avg_log2FC) %>% # Arrange the data.frame by decreasing avg_logFC ("-" = decreasing)
    group_by(cluster) %>% # Group the data.frame by cluster (this will produce a "grouped" data.frame, i.e. a subset for each cluster)
    top_n(N, avg_log2FC) %>% # Extract top N rows for each cluster based on avg_logFC
    arrange(cluster) 
}
#Old seu set
#Apply extractTopN() function
#DE_genes_SC_old_top12 <- extractTopN_old(DE_genes_SC_old, N = 12) #change avg_log2FC to old  avg_logFC in extractTopN
#geneSets_SC_old <- split(DE_genes_SC_old_top12$gene, paste0('',DE_genes_SC_old_top12$cluster))
#geneSets_SC_old

#geneSets_SC_old <- geneSets_SC_old[c(1:2,8:15,3:7)]
#geneSets_SC_old

#saveRDS(geneSets_SC_old, file = 'data/Article_SC_old_top12_15clusters.R')
geneSets_SC_old_article <- readRDS(paste0(path1, 'data/Article_SC_old_top12_15clusters.R'))
# new Seurat analysis, seu_2021_May 17 clusters SCT_snn_res.0.38 or 22 clusters incl muscle subclusters sub_cluster_SAN2
#seu <- readRDS('/Users/ChristerSylven/seu_2021_May.R')
#names(seu[[]])


#Idents(seu) <- "SCT_snn_res.0.38"
#DE_genes_SC_new  <- FindAllMarkers(seu, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
#dim(DE_genes_SC_new)

#new seu set & Muscle subset
#DE_genes_SC_new_top12 <- extractTopN(DE_genes_SC_new, N = 12) #change avg_log2FC to old  avg_logFC in extractTopN
#geneSets_SC_new <- split(DE_genes_SC_new_top12$gene, paste0('',DE_genes_SC_new_top12$cluster))
#geneSets_SC_new

# new large subset to get increasing order of clusters
#geneSets_SC_new <- geneSets_SC_new[c(1:2,10:17,3:9)]
geneSets_SC_new <- readRDS(paste0(path1, 'data/geneSets_SC_all_17_clusters.R'))


seu_Muscle <- readRDS(paste0(path1, 'seu_Muscle_2021_May.R'))
#names(seu_Muscle[[]])
Idents(seu_Muscle) <- "sub_cluster_SAN2"
DE_genes_SC_Muscle  <- FindAllMarkers(seu_Muscle, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
DE_genes_SC_Muscle_top12 <- extractTopN(DE_genes_SC_Muscle, N = 12) #change avg_log2FC to old  avg_logFC in extractTopN
geneSets_SC_Muscle<- split(DE_genes_SC_Muscle_top12$gene, paste0('',DE_genes_SC_Muscle_top12$cluster))
#geneSets_SC_Muscle

geneSets_SC_Muscle <- geneSets_SC_Muscle[1:8]
genes <- c("TBX18", "SHOX2", "HCN4") # sinus node
geneSets_SC_Muscle <- append(geneSets_SC_Muscle, list(genes))
names(geneSets_SC_Muscle)[9] <- 'SAN'
#geneSets_SC_Muscle


# cell rankings for whole ST_6 set
cells_rankings_ST_6 <- readRDS(paste0(path1, 'cells_rankings_ST_6.R'))
#cell rankings for spots present in both ST_6 and scMat_5 (Stereoscope)
cells_rankings_ST_6_subset <- cells_rankings_ST_6[,rownames(scMat_5)]
#dim(cells_rankings_ST_6_subset)

# AUC for old SC, 15 clusters, identical to Stereoscope
cells_AUC_SC_old_onST_article <- AUCell_calcAUC(geneSets_SC_old_article, cells_rankings_ST_6_subset, aucMaxRank = ceiling(0.05 * nrow(cells_rankings_ST_6_subset)), verbose = FALSE)

AUC_SC_old_onST_article <- setNames(data.frame(t(getAUC(cells_AUC_SC_old_onST_article))), nm = paste0('A', 1:nrow(getAUC(cells_AUC_SC_old_onST_article))-1))
#dim(AUC_SC_old_onST_article)

# AUC for new SC, 17 clusters
cells_AUC_SC_new_onST <- AUCell_calcAUC(geneSets_SC_new, cells_rankings_ST_6_subset, aucMaxRank = ceiling(0.05 * nrow(cells_rankings_ST_6_subset)), verbose = FALSE)

AUC_SC_new_onST <- setNames(data.frame(t(getAUC(cells_AUC_SC_new_onST))), nm = paste0('A', 1:nrow(getAUC(cells_AUC_SC_new_onST))-1))
#dim(AUC_SC_new_onST)

#
# AUC for new SC Muscle, 9 clusters
cells_AUC_SC_Muscle_onST <- AUCell_calcAUC(geneSets_SC_Muscle, cells_rankings_ST_6_subset, aucMaxRank = ceiling(0.05 * nrow(cells_rankings_ST_6_subset)), verbose = FALSE)

AUC_SC_Muscle_onST <- setNames(data.frame(t(getAUC(cells_AUC_SC_Muscle_onST))), nm = paste0('A', 1:nrow(getAUC(cells_AUC_SC_Muscle_onST))-1))
#dim(AUC_SC_Muscle_onST)


#Correlation
#corMat <- cor(AUC_ST_subset)

#corMat <- cor(scMat2)
#corMat <- cor(scMat)

#pheatmap::pheatmap(corMat, color = colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev())(51), breaks = seq(-1, 1, length.out = 51))

# Stereoscope vs SC old
sc <- t(AUC_SC_old_onST_article)
AUC <- t(scMat_5)
cormat1= matrix(0, nrow=nrow(sc), ncol=nrow(AUC))
for(i in 1:nrow(sc))
  {
    for(j in 1:nrow(AUC))
      {
           # cormat[i,j] = cor(unlist(sc[i,-1]), unlist(AUC[j,-1]))
      cormat1[i,j] = cor(sc[i,], AUC[j,])
        }
}
colnames(cormat1) = paste0("A", 1:ncol(cormat1)-1)
rownames(cormat1) = paste0("S", 1:nrow(cormat1)-1)

# Stereoscope vs Stereoscope
sc <- t(scMat_5)
AUC <- t(scMat_5)
cormat2= matrix(0, nrow=nrow(sc), ncol=nrow(AUC))
for(i in 1:nrow(sc))
{
  for(j in 1:nrow(AUC))
  {
    # cormat[i,j] = cor(unlist(sc[i,-1]), unlist(AUC[j,-1]))
    cormat2[i,j] = cor(sc[i,], AUC[j,])
  }
}
colnames(cormat2) = paste0("S", 1:ncol(cormat2)-1)
rownames(cormat2) = paste0("S", 1:nrow(cormat2)-1)
#c2 <- corrplot(cormat2, cl.pos='n') 

#AUC_SC_old_onST_article vs AUC_SC_old_onST_article
sc <- t(AUC_SC_old_onST_article)
AUC <- t(AUC_SC_old_onST_article)
cormat3= matrix(0, nrow=nrow(sc), ncol=nrow(AUC))
for(i in 1:nrow(sc))
{
  for(j in 1:nrow(AUC))
  {
    # cormat[i,j] = cor(unlist(sc[i,-1]), unlist(AUC[j,-1]))
    cormat3[i,j] = cor(sc[i,], AUC[j,])
  }
}
colnames(cormat3) <- colnames(AUC_SC_old_onST_article)
rownames(cormat3) <- colnames(AUC_SC_old_onST_article)

#AUC_SC_new_onST vs AUC_SC_new_onST
sc <- t(AUC_SC_new_onST)
AUC <- t(AUC_SC_new_onST)
cormat4= matrix(0, nrow=nrow(sc), ncol=nrow(AUC))
for(i in 1:nrow(sc))
{
  for(j in 1:nrow(AUC))
  {
    # cormat[i,j] = cor(unlist(sc[i,-1]), unlist(AUC[j,-1]))
    cormat4[i,j] = cor(sc[i,], AUC[j,])
  }
}
colnames(cormat4) <- colnames(AUC_SC_new_onST)
rownames(cormat4) <- colnames(AUC_SC_new_onST)

#AUC_SC_Muscle_onST_article vs AUC_SC_oMuscle_onST_article
sc <- t(AUC_SC_Muscle_onST)
AUC <- t(AUC_SC_Muscle_onST)
cormat5= matrix(0, nrow=nrow(sc), ncol=nrow(AUC))
for(i in 1:nrow(sc))
{
  for(j in 1:nrow(AUC))
  {
    # cormat[i,j] = cor(unlist(sc[i,-1]), unlist(AUC[j,-1]))
    cormat5[i,j] = cor(sc[i,], AUC[j,])
  }
}

colnames(AUC_SC_Muscle_onST) <- c('Trabecular', 'Compact', 'Purkinje-related',' Atrial trabecular', 'HighG2M','Atrial conduit', 'Exosome-related', 'OFT', 'SAN')

colnames(cormat5) <- colnames(AUC_SC_Muscle_onST)
rownames(cormat5) <-   c('','','','','','','','','')


nf <- layout(matrix(c(1,2,3,4,5,6), 2, 3, byrow = TRUE), widths = c(0.75, 0.75, 1), respect = TRUE)
#layout.show(nf)

#par(mfcol=c(2,3))

corrplot(cormat1, cl.pos='n')
corrplot(cormat2, cl.pos='n')

corrplot(cormat5, cl.pos='n', tl.cex = 1, tl.srt =90) 
res1 <- cor.mtest(cormat5, conf.level = .99)
## specialized the insignificant value according to the significant level
corrplot(cormat5,  cl.pos='n', add = TRUE, p.mat = res1$p, sig.level = .01,  type = 'lower', tl.pos = "n")
corrplot(cormat3, cl.pos='n')
corrplot(cormat4, cl.pos='n')

```

## Session info

```{r}
sessionInfo()
```
